### 2020-03-25 engineering.tex
spdl description, use case and requirements draft

\epigraph{The difference between ideas and reality is like the
difference between philosophy and engineering. The work to transform one into
the other is scientific research}{{\itshape V-Research}}

Several standards mandates a secure-by-design approach in which 
cybersecurity shall be considered at the very early stages of the 
design process. For example,
\begin{enumerate}[noitemsep]
	\item DO-326A -- ``Airworthiness Security Process Specification''
		requires a cybersecurity risk assessment of the design and is
		the ``are the only Acceptable Means of Compliance (AMC) by FAA
		\& EASA for aviation cyber-security airworthiness certification,
		as of 2019'' as pointed out by SAE in \autocite{SAE2019DO326A}.
	\item NIST 800-82 \autocite{Stouffer2011guide} -- ``guide to Industrial
		Control System (ICS) Security''
	\item J3061:2016-1 \autocite{SAE2016J3061} -- ``Cybersecurity Guidebook
		for Cyber-Physical Vehicle Systems'' defines ``set of
		high-level guiding principles for Cybersecurity as it relates
		to cyber-physical vehicle systems'' and states that
		``incorporate Cybersecurity into cyber-physical vehicle systems
		from concept phase through production, operation, service, and
		decommissioning''
\end{enumerate}
Companies who develop CPS (e.g. aerospace systems or elevator systems), must adhere
to strict regulations and quality assurances processes. Introducing tools and procedures 
in their processes requires a detailed and justified overview on how those
tools and procedures can be used. In Figure~\ref{fig:spdl}, we provide an overview
of a process that consider cybersecurity in the specification, design, and implementation
stages. The overall process start with the specification of the architectural (both physical
and functional) requirements from which the Weaknesses are automatically identified.
Based on the Assets, architectures, and the Weaknesses the cybersecurity risk is
calculated based on the $\abf$-theory. An iterative process between 
the user (e.g. the engineer) and the \secramod module of the \abftool allows the user to move to the design phase
when the risk level is considered acceptable. The design phase starts by automatically mapping
the Weaknesses into Vulnerabilities which are, in turn, fed into the \designverifmod
module along with the final specification. The Module returns a list of Vulnerabilities and
potential Mitigations. The user then procees in implementing the CPS and the SW/HW choices
can be tested by the \atgmod.

\begin{figure}[t]
	\centering
	\includegraphics[width=\textwidth]{spdl.pdf}
	\caption{Secure-by-design System Development Life-cycle}
	\label{fig:spdl}
\end{figure}

In order to explain in details the SPDL, we use the following running example.
\begin{example}
	As depicted in Figure~\ref{fig:usecase}, the use case of the running example
	(a small part of a subprocess of the SwAT testbed \autocite{}), shows
	\begin{itemize}[noitemsep]
		\item a tank that is filled with raw water coming from the inlet pipe 
		\item a motorized valve (a valve with an actuator) that regulates that
			\begin{itemize}
				\item if opened, allows the raw water to flow into the tank
				\item if closed, stops the raw water to flow into the tank
			\end{itemize}
		\item a sensor that reads the meters of raw water in the tank
		\item a PLC controller, connected to the sensor and the actuator, that receives the readings from the sensor and regulates the quantity of water in the tank by communicating a change in the state of the actuator based on the following logi
			\begin{itemize}
				\item if the readings from the sensor reports a water level in the tank above (or equal) to 10m, the actuator shall close the valve
				\item if th readings are below 10m the actuator shall open the valve.
			\end{itemize}
	\end{itemize}

\begin{figure}[t]
	\centering
	\includegraphics[width=.55\textwidth]{usecase.pdf}
	\caption{Use Case -- Running Example}
	\label{fig:usecase}
\end{figure}
\end{example}

\subsection{Specification -- Elicitation of Security Requirements}\label{sec:properties}
\begin{figure}[t]
	\centering
	\includegraphics[width=\textwidth]{spec.pdf}
	\caption{Specification -- Running Example}
	\label{fig:spec}
\end{figure}
The first step of the SPDL is the definition of the requirements and then, for
the purpose of this paper of the (i) security requirements, (ii) architectural
(physical and functional) requirements.  Our focus in mainly on security
requirements but they are necessarily intertwined with the architectural
requirements, as we will show afterwards in this section.  A complete study on
how to properly define the architectural requirements is outside the scope of
this paper so we propose a method that is convenient for the purpose of our
assessment.

\Paragraph{System Requirements}
In our running example, there is a system (a CPS) composed by $5$ main
subsystem which we consider atomic (i.e. not themself composed by subsystems)
and then agents: the tank, the sensor, the controller, the actuator, and the
motorized valve (valve from now on).  The whole specification (which we now
describe) of the physical and functional architecture is depicted in
Figure~\ref{fig:spec}.  In the use case, the communication flow is as follows:
\begin{enumerate}
	\item the sensor reads the level of water in the tank ($\text{tank}\rightarrow \text{sensor}$) \footnote{this communication depends on the technology in use for the readings. We assume a unidirectional communication without loss of generality}
	\item the sensor communicates the readings to the controller ($\text{sensor}\rightarrow \text{controller}$)
	\item the controller calculates the valve state and communicates it to the actuator ($\text{controller}\rightarrow \text{actuator}$)
	\item the actuator translates (digital to analog) the states received and communicates\footnote{we do not distinguish, for the sake of simplicity between different types of channels and assume that communicating to the valve produces the expected change of physical state} it to the valve
\end{enumerate}
Each agent has an input or output port from which it communicates with other
agents.  It is also composed by the functional blocks that defines its
functional architecture and physical boundaries that defines its physical
architecture, as follows.
\begin{enumerate}[start=1, label={req\arabic*)}]
	\item the tank has no functional architecture, it is a purely physical component which has the only purpose of containing water
	\item the communication is unidirectional from the tank to the input port of the sensor
	\item the sensor has an input-port that receives only incoming communication from the tank
	\item the sensor has two functional blocks
		\begin{enumerate}[label={req4.\arabic*)}]
			\item read\_water\_level, accepts inputs only from the input-port and outputs a new reading every $t$ seconds
			\item comm\_water\_level, accepts inputs only from the read\_water\_level and communicates the readings to the output-port
		\end{enumerate}
	\item the sensor has an output-port that only outputs the messages generated by the comm\_water\_level to the input-port of the controller
	\item the controller has in input-port that only receives incoming communications from the sensor and sends the messages to the functional architecture
	\item the functional architecture of the controller is composed by three functional blocks
		\begin{enumerate}[label={req7.\arabic*)}]
			\item get\_water\_level, accepts inputs only from the input-port and outputs the water level to the main control function
			\item control, calculate the next state of the valve based on the inputs receive from the get\_water\_level
			\item send\_state, receives the output of the control function and communicates it to the output-port
		\end{enumerate}
	\item the controller has an output-port that only outputs the messages generated by the send\_state to the input-port of the actuator 
	\item the actuator has an input-port that receives only incoming communication from the controller 
	\item the functional architecture of the actuator is composed by two functional blocks
		\begin{enumerate}[label={req10.\arabic*)}]
			\item read\_change\_state, accepts inputs only from the input-port and outputs the next state of the valve to the close\_open\_valve function
			\item close\_open\_valve, converts the digital next state received from the read\_change\_state an analog signal and communicates it the output-port
		\end{enumerate}
	\item the valve accepts analog messages from the output port of the actuator and changes its state based on those messages
\end{enumerate}

While \emph{agents} and \emph{communication} between agents are described and
defined in Section~\ref{sec:theory}, both the concepts of \emph{port},
\emph{channel}, \emph{functional and physical architecture}, and
\emph{functional block} have not yet been defined with respect to the $\abf$-theory.
Informally, a channel identifies the place where the communication takes place.
In this work we only consider mono-directional channels and communication but the 
extension to bi-directional channel can be considered as the union of two
unidirectional channels.

\begin{definition}{\bf mono-directional channel --}\label{def:monochannel}
	A mono-directional channel is defined by the message sent and received
	over the channel. Given two agents $s$ (sender) and $r$ (receiver) a
	mono-directional channel $s \rightarrow r$ defines the sending of
	information between $s$ and $r$. 
	%$\assertion{s}{r}$ is defined as a (improper) subset of the assertions of $s$, i.e.
	%$\pp{\assertion{s}{r}}{\assertion{s}}
	%PP(A λ S ,λ R , A λ S ) ∨ EQ(A λ S ,λ R , A λ S ) (since λ S may
	%communicate with other agents).  The receipt of messages over a
	%mono-directional channel is defined as the topolog- ical boundary B S
	%,R (improper) subset of the beliefs of the receiver (i.e., the beliefs
	%generated by the message exchange and by computations over the set of
	%beliefs): PP(B S ,R , B R ) ∨ EQ(B S ,R , B R ).
\end{definition}

\begin{definition}{\bf \fix{mr}{informal Input or Output Port} --}\label{def:port} 
	Ports are agents with the following predefined behavior: an input-port
	changes the type of information from assertion to belief while an
	output-port from belief to assertion; forwarding information from the
	outside of an agent's boundary to the inside (input-port) or vice versa
	(output-port).
\end{definition}

\Paragraph{Security Requirements} 
While standards (such
as IEC62443-1-3\fixnote{mr}{check the correct version}) defines requirements as
``information in transit/at-rest should be encrypted'' we believe that, at
specification level, security requirements can be formulated as subcategories of
the well-known CIA triad (where CIA stands for Confidentiality, Integrity,
Availability).
Infosec, or Information Security, defines the security risks related to
information with the de-facto standard CIA triad , which is often criticized
\autocite{CIAcriticismCPS} for being too general or non-adequate (e.g. by
adding authenticity which is often used as a building block for confidentiality
and integrity) to be effectively applied to the engineering of systems. Due to
this, many researchers an organization tried to improve the CIA triad.  The
evolutions/extensions of the CIA triad has been documented, e.g., in
\autocite{Samonas2014cia} and is summarized in Table~\ref{tab:ciaevolution}.

\begin{table}[h]
\centering
%\setlength{\tabcolsep}{3.5pt}
%\renewcommand{\arraystretch}{1}
\small
\begin{tabular}{rll} 
	{\bf Year} & {\bf Definition} & {\bf Legend}\\
	\hline
	1970s & infosec = CIA & Confidentiality, Integrity, Availability\\
	1980s & infosec += (Au, nR) & Authenticity and non-Repudiation\\
	1990s & infosec += CSpec & Correctness in Specification\\
	2000s & infosec += RITE & Responsibility, Integrity of people, Trust, Ethicality
\end{tabular}
\caption{Chronological progression of the CIA triad~\label{tab:ciaevolution}}
\end{table}

An representative example is the effort made by the OECD (Organisation for Economic
Co-operation and Development) in \autocite{OECD2013guidelines} to define
security guidelines for information system (1992) based on new principles such
as ``awareness, ethics, risk assessment'' and maintain those revising the
document (e.g.  following the multistakeholder expert consultation in 2013).
Other similar efforts focused on defining entirely new principles or extending
the CIA triad, such as \autocite{NISTSP800-160}. What is of interest for our
argument is that, to the best of our knowledge, all of the approaches that aim
at improving the CIA triad are besed on motivations related to empirical evidences. On the contrary,
we now define (in Section~\ref{sec:properties}) the CIA triad for an agent
defined in the $\abftheory$ theory and we test if (i) other properties are
allowed in the $\abftheory$ theory, and (ii) if and how the CIA triad can be
detailed in the $\abftheory$ theory.

In \autocite{Anderson1972report,Samonas2014cia}, CIA are defined as follows.
\begin{itemize}
	\item Unauthorized information release: an unauthorized person is able
		to read and take  advantage  of  information  stored  in  the
		computer.  This  category of concern  sometimes  extends  to
		``traffic  analysis,''  in  which  the intruder  only observes
		the  patterns  of  information  use.  From  those patterns,
		the  intruder can   infer   some   information   content.
		This category   also   includes   the unauthorized use of a
		proprietary program.  (Confidentiality or Secrecy) 
	\item Unauthorized  information  modification:  an  unauthorized person
		is  able  to make changes in stored information -- a form  of
		sabotage.  It should be noted that in the case of this kind of
		violation, the intruder does not necessarily see the
		information he has changed.  (Integrity)
	\item Unauthorized  denial  of  use:  an  intruder  can  prevent  an
		authorized  user  from referring to, or from modifying
		information, even though the intruder may not be able to refer
		to, neither modify the information themselves. (Availability)
\end{itemize}

We now want to generalize the CIA triad to agents and not just 
information and map it to the $\abftheory$ theory.
\begin{definition}{\bf Infosec-Confidentiality --}\label{def:infosec-c}
	An Information $\varphi$ is considered to be confidential to a set of 
	agents $Ag_c$ if, per each agent $a\in Ag_c$, the content of the
	information (i.e. the data it carries) can only be understood (i.e. the relation
	between the information and its data can be known) by the
	agents in $Ag_c$.

	\begin{itemize}
		\item[$(\interpretation21)$] for all agents $a\in Ag$,
			$\world\models\information{a}{\varphi}$ and
			$\world\models\rcc(\varphi,Conf) \wedge
			\neg\dr{Conf}{\varphi}$ iff for any world $\world'$
			such that $\world \modalrelation\world'$,
			$\world'\models\believe{a}{\varphi}$, and
			$\world'\models\knows{a}{\exists f.~
			f(\information{a}{\varphi})=\varphi}$, where 
			$Conf$ is a Region and $f$ is an uninterpreted function.
	\end{itemize}
\end{definition}

\begin{example}{Dummy security protocol --}
	The last part (part III) of the protocol in our first example
	(Figure~\ref{fig:protocol-example}) represents the actual key exchange
	which, in general, needs to be confidential; otherwise the key would be
	known by any other agent who can, e.g., read on the communication
	channel.  Therefore, the last message, in Alice-and-Bob notation can be
	rephrased as: \begin{displaymath} Alice\Rightarrow Bob: enc(pk(Bob),
	symmetricKey) \end{displaymath} where $enc$ is the standard asymmetric
	encryption operator that encrypts the message $symmetricKey$ with the
	public key of Bob $pk(Bob)$ (see, e.g.,
	\autocite{Rocchetto2017interpolation} for a detailed formalization).
	The function $f$ in Definition~\ref{def:infosec-c}, represents any
	function used to enforce confidentiality on information.  In this
	example, $f$ is identified with $enc$ (i.e.
	$enc(pvt(Bob),symmetricKey)=symmetricKey$), the information is
	$\information{Alice}{[enc(pk(Bob),symmetricKey)]}$ and the data, i.e.
	$\varphi$ in Definition~\ref{def:infosec-c}, is $\varphi=symmetricKey$.
	In the perfect cryptography assumption, and assuming for simplicity
	that the logic of the whole protocol preserves confidentiality, the
	message is confidential. In fact, if $Ag={Alice, Bob}$ and $\eq{symmetricKey}{Conf}$:
	\begin{align*}
		\world^0\models\belief{Alice}{symmetricKey},\\
		\world^1\models\information{Alice}{symmetricKey}, \text{and}\\
		\world^2\models\belief{Bob}{symmetricKey} ~\text{if} \\
		enc(pvt(Bob), \information{Alice}{symmetricKey}=symmetricKey)=symmetricKey\\
		(i.e. f=enc ~\text{and}~ \information{Alice}{symmetricKey}=enc(pub(Bob),symmetricKey))\\
	\end{align*}
\end{example}

\begin{definition}{\bf Agent-Confidentiality --}\label{def:confidentiality}
An agent preserves the property of confidentiality iff 
\end{definition}

\begin{definition}{\bf System-Confidentiality --}\label{def:confidentiality}
	The property of confidentiality holds in a system when its information
	preserves confidentiality.
\end{definition}

\fix{mr}{
Confidentiality: the information is accessible only to authorised individuals.
A set of rules that limits access to information.  The information can be
accessed by X if and only if X is authorized to access it = X has the right
attributes (i.e. keys, password) to access it.

Integrity: the information is trustworthy and accurate (i.e. uncorrupted and
unaltered). Protects data from modification or deletion by unauthorized sources
and the ability to undo any damage done.  (for data at rest) If the information
X is written, A will always be read. If information X is modified to
information X', information X can always be recovered.  (for data in transit)
If the information X is sent, A will be received. If information X is modified
to information X', information X can always be recovered.

Availability: the information is accessible and usable when required. Guarantee
of reliable access to the information by authorized people If the information X
needs to be accessed, information X can be retrieved. If

Authenticity: assurance that the information is from the source it claims to be
from. Authenticity involves proof of identity. Authenticity is verified through
authentication.  If information X arrives from A to B, A sent information X and
B can verify it.  ??Authenticity implies integrity but Integrity doesn't imply
authenticity??  ??Authenticity doesn't imply confidentiality but
confidentiality does imply authenticity??

Non-repudiation: the ability to ensure that a party to a contract or a
communication cannot deny the authenticity of their signature on information
(or the sending of) that they originated.  If information X arrives from A, A
sent it and cannot deny it (partial overlap??).  If sender A sends information
X, it cannot say he didn't do it.  Non-repudiation implies authenticity and
integrity. Authenticity doesn't imply non-repudiation (replay attacks?).
Integrity doesn't imply non-repudiation (replay attacks?).

\begin{enumerate}
\item specification -- definition of the desired design of a system. The
specification is verified w.r.t. the  $\abf$ theory. Checking the
controls of the whole cybersecurity life-cycle and the relation with
CWE.
\item design -- the mitigations identified in the specification stage
are implemented into the design. The security of the assets w.r.t. the design are
verified with a, so called, cybersecurity risk assessment.
\end{enumerate}
}