### 2020-03-30 insecureport.tex
first complete draft communication insecurity

\fix{mr}{this proof needs to be done}
\begin{proof}
In Appendix~\ref{app:commproof}, we report the proof which is exhaustive on all
possible combinations.  In the following we report only the excluded cases and
assume the other to be strightforwad (as a proof sketch).
\begin{enumerate}
	\item \emph{input secure port} and 
		\begin{enumerate}
			\item \emph{output drop port} ($s\sdcom r$), 
			\item \emph{output insertion port} ($s\sicom r$), 
			\item \emph{output injection port} ($s\sjcom r$), 
			\item \emph{output selective drop port} ($s\ssdcom r$), 
			\item \emph{output selective insertion port} ($s\ssicom r$), 
			\item \emph{output selective injection port} ($s\ssjcom r$), 
		\end{enumerate}
	\item \emph{input drop port} (all the input-assertions are dropped) and 
		\begin{enumerate}
			\item \emph{output drop port} ($s\ddcom r$), this
				configuration is possible but the input port
				doesn't forward any message therefore the
				output port will never have a message to drop
			\item \emph{output insertion port} ($s\dicom r$), where
				all the ouput-assertions have been inserted by
				the output port
			\item\label{drop-injection} \emph{output injection
				port}, this configuration is impossible since
				there is no Assertions from the input-port to
				inject (or it behaves as $s\dicom r$)
			\item\label{drop-seldrop} \emph{output selective drop
				port}, as in \ref{drop-injection} there is no
				Assertion from the input port therfore no
				selective port is possible in general. In this
				case this configuration is impossible or
				behaves as $s\ddcom r$
			\item \emph{output selective insertion port}, as in
				\ref{drop-seldrop} impossible, or behaves as
				$s\dicom r$
			\item \emph{output selective injection port}, as in
				\ref{drop-seldrop} impossible, or behaves as
				$s\dicom r$
			\item \emph{ouput secure port} ($s\dscom r$)
		\end{enumerate}

	\item \emph{input insertion port} (where all the assertion reaching the
		output-port have been artificially generated by the input-port) and 
		\begin{enumerate}
			\item \emph{output drop port} ($s\idcom r$), this
				configuration is possible but the drop port
				drops every inserted message.
			\item \emph{output insertion port} ($s\iicom r$), new
				Assertions have all been inserted both at the left-hand
				side and righ-hand side of the communication.
			\item \emph{output injection port} ($s\ijcom r$), the
				new Assertions, inserted in the left-hand side,
				are then modified/injected on the right-hand side of the
				communication
			\item \emph{output selective drop port} ($s\isdcom r$),
				where some insertions of the input port are dropped 
			\item \emph{output selective insertion port} ($s\isicom
				r$), where some insertions of the input port
				are substituted with new insertion on the
				right-hand side of the communication
			\item \emph{output selective injection port} ($s\isjcom
				r$) where some insertions of the input port are
				modified/injected on the right-hand side of the
				communication 
			\item \emph{output secure port} ($s\iscom r$)
		\end{enumerate}

	\item \emph{input injection port} and 
		\begin{enumerate}
			\item \emph{output drop port} ($s\jdcom r$)
			\item \emph{output insertion port} ($s\jicom r$)
			\item \emph{output injection port} ($s\jjcom r$)
			\item \emph{output selective drop port} ($s\jsdcom r$)
			\item \emph{output selective insertion port} ($s\jsicom r$)
			\item \emph{output selective injection port} ($s\jsjcom r$)
			\item \emph{output secure port} ($s\jscom r$)
		\end{enumerate}

	\item \emph{input selective drop port} and 
		\begin{enumerate}
			\item \emph{output drop port} ($s\sddcom r$)
			\item \emph{output insertion port} ($s\sdicom r$)
			\item \emph{output injection port} ($s\sdjcom r$)
			\item \emph{output selective drop port} ($s\sdsdcom r$)
			\item \emph{output selective insertion port} ($s\sdsicom r$)
			\item \emph{output selective injection port} ($s\sdsjcom r$)
			\item \emph{output secure port} ($s\sdscom r$)
		\end{enumerate}

	\item \emph{input selective insertion port} and 
		\begin{enumerate}
			\item \emph{output drop port} ($s\sidcom r$)
			\item \emph{output insertion port} ($s\siicom r$)
			\item \emph{output injection port} ($s\sijcom r$)
			\item \emph{output selective drop port} ($s\sisdcom r$)
			\item \emph{output selective insertion port} ($s\sisicom r$)
			\item \emph{output selective injection port} ($s\sisjcom r$)
			\item \emph{output secure port} ($s\siscom r$)
		\end{enumerate}

	\item \emph{input selective injection port} and 
		\begin{enumerate}
			\item \emph{output drop port} ($s\sjdcom r$)
			\item \emph{output insertion port} ($s\sjicom r$)
			\item \emph{output injection port} ($s\sjjcom r$)
			\item \emph{output selective drop port} ($s\sjsdcom r$)
			\item \emph{output selective insertion port} ($s\sjsicom r$)
			\item \emph{output selective injection port} ($s\sjsjcom r$)
			\item \emph{output secure port}  ($s\sjcom r$)
		\end{enumerate}
\end{enumerate}
\end{proof}