### 2020-04-16 tool.tex
add modelio project final draft and XMI parser

\begin{itemize}
\item structured in package where the \emph{architecture} is
\begin{itemize}
\item is defined by an object diagram where the \emph{physical architecture} is structured on dependency connections and
\item the \emph{functional architecture} is linked to the agents and defined as nested/inner object diagrams
\item the information flow between functional or physical components is explicit 
\end{itemize}
\end{itemize}