### 2020-06-08 reports/conference_paper/main.tex
first full draft of the conference paper

\documentclass[conference]{IEEEtran}
\usepackage[table]{xcolor}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[style=numeric]{biblatex}
\usepackage{amsmath}
\addbibresource{../literature.bib}% Syntax for version >= 1.2
\usepackage{graphicx}
\graphicspath{{../images/}}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{amsthm}
\usepackage{rotating}

% circles
\usepackage{wasysym}
\newcommand{\Tdot}{$\CIRCLE$}
\newcommand{\Thdot}{$\LEFTcircle$}
\newcommand{\Twdot}{$\Circle$}

% for fancy table with rotated column titles
\usepackage{tabularx}
\usepackage{adjustbox}
\newcolumntype{R}[2]{%
    >{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
    l%
    <{\egroup}%
}
\newcommand*\rot{\multicolumn{1}{R{45}{0.01em}}}% no optional argument here, please!
\newcommand*\rota{\multicolumn{1}{R{90}{0.01em}}}% no optional argument here, please!
\def\hb{\hbox to 10.7 cm{}}

%colors in tables
\definecolor{abfred}{RGB}{255, 215, 215}
\definecolor{abf-rg-1}{RGB}{251,213,242}
\definecolor{abf-rg-2}{RGB}{243,212,249}
\definecolor{abf-rg-3}{RGB}{213,210,245}
\definecolor{abf-rg-4}{RGB}{208,231,242}
\definecolor{abf-rg-5}{RGB}{206,238,223}
\definecolor{abf-rg-6}{RGB}{205,236,210}
\definecolor{abf-rg-7}{RGB}{211,234,204}
\definecolor{abfgreen}{RGB}{221,232,203}
\definecolor{abforange}{RGB}{255, 245, 206}

%fix commands
\newcommand{\fixnote}[2]{\textbf{\color{red}{FIX}}\footnote{{\bf #1:} #2}}
\newcommand{\fix}[2]{{\color{red} {\bf tofix:} #2}}
% \renewcommand{\fixnote}[2]{}
% \renewcommand{\fix}[2]{}

\newcommand{\assertionRegion}{\mathcal{A}}
\newcommand{\beliefRegion}{\mathcal{B}}
\newcommand{\factRegion}{\mathcal{F}}
\newcommand{\rcc}{rcc}
\newcommand{\agentuniverse}{Ag}
\newcommand{\abf}{\assertionRegion,\beliefRegion,\factRegion}
\newcommand{\abftheory}{\assertionRegion\beliefRegion\factRegion}
\newcommand{\Rcc}[2]{rcc(#1,#2)}
\newcommand{\connects}[2]{C(#1,#2)}
\newcommand{\disconnected}[2]{\neg C(#1,#2)}
\newcommand{\partof}[2]{P(#1,#2)}
\newcommand{\overlaps}[2]{O(#1,#2)}
\newcommand{\eq}[2]{EQ(#1,#2)}
\newcommand{\pp}[2]{PP(#1,#2)}
\newcommand{\po}[2]{PO(#1,#2)}
\newcommand{\ppi}[2]{PPi(#1,#2)}
\newcommand{\dr}[2]{DR(#1,#2)}
\newcommand{\all}[2]{T(#1,#2)}
\newcommand{\rassert}[3]{\mathcal{A}_{#1\rightarrow #2}#3}
\newcommand{\lassert}[3]{\mathcal{A}_{#1\leftarrow #2}#3}

\newcommand*\halfbullet[1][0.4ex]{%
  \begin{tikzpicture}
	  \raisebox{0.6pt}{\draw[fill] (0,0)-- (90:#1) arc (90:270:#1) -- cycle ;
	  \draw (0,0) circle (#1);}
  \end{tikzpicture}}

\newcommand{\fakedmapsfromchar}{%
  \mathrel{\mathpalette\reflectmathchar\mapstochar}%
}
\makeatletter
\newcommand{\reflectmathchar}[2]{%
  \reflectbox{$\m@th#1#2$}%
}
\makeatother
\newcommand{\rightbararrow}{\rightarrow\fakedmapsfromchar}
\newcommand{\leftbararrow}{\mapstochar\rightarrow}
\newcommand{\rightleftbararrow}{\mapstochar\rightarrow\fakedmapsfromchar}
\setlength{\fboxrule}{0.1pt}
\setlength{\fboxsep}{1.5pt}
\newcommand{\secureportelem}{\mathrel{\rightarrow}}
\newcommand{\secureport}{\fbox{\ensuremath \secureportelem}}
\newcommand{\dropportelem}{\mathrel{\circ\!\!\rightarrow}}
\newcommand{\dropport}{\fbox{\ensuremath \dropportelem}}
\newcommand{\insertportelem}{\mathrel{\bullet\!\!\rightarrow}}
\newcommand{\insertport}{\fbox{\ensuremath \insertportelem}}
\newcommand{\injectionportelem}{\mathrel{\halfbullet\!\!\rightarrow}}
\newcommand{\injectionport}{\fbox{\ensuremath \injectionportelem}}
\newcommand{\selectiveportelem}{\mathrel{\leftbararrow}}
\newcommand{\selectiveport}{\fbox{\ensuremath \selectiveportelem}}
\newcommand{\selectiveinsertportelem}{\mathrel{\bullet\!\!\leftbararrow}}
\newcommand{\selectiveinsertport}{\fbox{\ensuremath \selectiveinsertportelem}}
\newcommand{\selectiveinjectionportelem}{\mathrel{\halfbullet\!\leftbararrow}}
\newcommand{\selectiveinjectionport}{\fbox{\ensuremath \selectiveinjectionportelem}}
\newcommand{\selectivedropportelem}{\mathrel{\circ\!\!\leftbararrow}}
\newcommand{\selectivedropport}{\fbox{\ensuremath \selectivedropportelem}}
%communication channels
%secure input port
\newcommand{\sdcom}{\mathrel{\rightarrow\!\!\circ}}
\newcommand{\sicom}{\mathrel{\rightarrow\!\!\bullet}}
\newcommand{\sjcom}{\mathrel{\rightarrow\!\!\halfbullet}}
\newcommand{\ssdcom}{\mathrel{\rightbararrow\!\!\circ}}
\newcommand{\ssicom}{\mathrel{\rightbararrow\!\!\bullet}}
\newcommand{\ssjcom}{\mathrel{\rightbararrow\!\halfbullet}}
%drop input port
\newcommand{\ddcom}{\mathrel{\circ\!\!\rightarrow\!\!\circ}}
\newcommand{\dicom}{\mathrel{\circ\!\!\rightarrow\!\!\bullet}}
\newcommand{\dscom}{\mathrel{\circ\!\!\rightarrow}}
%insert input port
\newcommand{\idcom}{\mathrel{\bullet\!\!\rightarrow\!\!\circ}}
\newcommand{\iicom}{\mathrel{\bullet\!\!\rightarrow\!\!\bullet}}
\newcommand{\ijcom}{\mathrel{\bullet\!\!\rightarrow\!\!\halfbullet}}
\newcommand{\isdcom}{\mathrel{\bullet\!\!\rightbararrow\!\!\circ}}
\newcommand{\isicom}{\mathrel{\bullet\!\!\rightbararrow\!\!\bullet}}
\newcommand{\isjcom}{\mathrel{\bullet\!\!\rightbararrow\!\halfbullet}}
\newcommand{\iscom}{\mathrel{\bullet\!\!\rightarrow}}
%injection input
\newcommand{\jdcom}{\mathrel{\halfbullet\!\!\rightarrow\!\!\circ}}
\newcommand{\jicom}{\mathrel{\halfbullet\!\!\rightarrow\!\!\bullet}}
\newcommand{\jjcom}{\mathrel{\halfbullet\!\!\rightarrow\!\!\halfbullet}}
\newcommand{\jsdcom}{\mathrel{\halfbullet\!\!\rightbararrow\!\!\circ}}
\newcommand{\jsicom}{\mathrel{\halfbullet\!\!\rightbararrow\!\!\bullet}}
\newcommand{\jsjcom}{\mathrel{\halfbullet\!\!\rightbararrow\!\halfbullet}}
\newcommand{\jscom}{\mathrel{\halfbullet\!\!\rightarrow}}
%selective drop input
\newcommand{\sddcom}{\mathrel{\circ\!\!\leftbararrow\!\!\circ}}
\newcommand{\sdicom}{\mathrel{\circ\!\!\leftbararrow\!\!\bullet}}
\newcommand{\sdjcom}{\mathrel{\circ\!\!\leftbararrow\!\!\halfbullet}}
\newcommand{\sdsdcom}{\mathrel{\circ\!\!\rightleftbararrow\!\!\circ}}
\newcommand{\sdsicom}{\mathrel{\circ\!\!\rightleftbararrow\!\!\bullet}}
\newcommand{\sdsjcom}{\mathrel{\circ\!\!\rightleftbararrow\!\halfbullet}}
\newcommand{\sdscom}{\mathrel{\circ\!\!\leftbararrow}}
%selective insertion
\newcommand{\sidcom}{\mathrel{\bullet\!\!\leftbararrow\!\!\circ}}
\newcommand{\siicom}{\mathrel{\bullet\!\!\leftbararrow\!\!\bullet}}
\newcommand{\sijcom}{\mathrel{\bullet\!\!\leftbararrow\!\!\halfbullet}}
\newcommand{\sisdcom}{\mathrel{\bullet\!\!\rightleftbararrow\!\!\circ}}
\newcommand{\sisicom}{\mathrel{\bullet\!\!\rightleftbararrow\!\!\bullet}}
\newcommand{\sisjcom}{\mathrel{\bullet\!\!\rightleftbararrow\!\halfbullet}}
\newcommand{\siscom}{\mathrel{\bullet\!\!\leftbararrow}}
%selective injection
\newcommand{\sjdcom}{\mathrel{\halfbullet\!\!\leftbararrow\!\!\circ}}
\newcommand{\sjicom}{\mathrel{\halfbullet\!\!\leftbararrow\!\!\bullet}}
\newcommand{\sjjcom}{\mathrel{\halfbullet\!\!\leftbararrow\!\!\halfbullet}}
\newcommand{\sjsdcom}{\mathrel{\halfbullet\!\!\rightleftbararrow\!\!\circ}}
\newcommand{\sjsicom}{\mathrel{\halfbullet\!\!\rightleftbararrow\!\!\bullet}}
\newcommand{\sjsjcom}{\mathrel{\halfbullet\!\!\rightleftbararrow\!\halfbullet}}
\newcommand{\sjscom}{\mathrel{\halfbullet\!\!\leftbararrow}}

\newtheorem{definition}{Definition}%[section]
\newtheorem{hypothesis}{Hypothesis}%[section]

\begin{document}
\title{On Etiology of Cybersecurity}

\author{\IEEEauthorblockN{Marco Rocchetto}
\IEEEauthorblockA{V-Research\\
Verona, Italy\\
Email: marco@v-research.it}
\and
\IEEEauthorblockN{Francesco Beltramini}
\IEEEauthorblockA{V-Research\\
Verona, Italy\\
Email: francesco@v-research.it}
}

\maketitle

\begin{abstract}
The objective of this research is the development of a theory that defines (all
	and only) the possible insecurity and security configurations of any
	abstract system. The theory is structured upon other theories that
	defines how a component of a system can be abstracted into an agent,
	defining how agents can be formalized (both syntactically and
	semantically) to describe an abstract system, such as a graph. The core
	theories are the epistemological definition of knowledge, Beliefs, and
	Information, the Assertion-Belief-Fact framework, and mereotopology.
	We implemented a formal theory (of axioms) of a mereotopology, and of
	the Region Connection Calculus (RCC3 and RCC5) in a Python program that
	uses the Z3 SMT solver. The results show that a single component (i.e.
	agent) of an abstract system has a definite number of  different
	insecurity configurations (e.g. 53 using RCC5 over a topological
	structure) and only 1 secure (i.e.  expected) configuration. The
	configurations are reported as models satisfying the abstract system
	semantics. We use these results to predict all possible security
	weaknesses of a system, instead of relying on databases of known
	weaknesses, and to lay the foundations of a scientific theory of
	cybersecurity.  We show a concrete applications of our theory to the
	risk assessment of an ad-hoc system.
\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction}\label{sec:intro}
In \cite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
``an asymmetry in computer security'', which he defines as follows: ``Things
can be declared insecure by observation, but not the reverse. There is no
observation that allows us to declare an arbitrary system or technique
secure''. Herley then uses this argument to show that ``claims that any measure
is necessary for security are empirically unfalsifiable''. Given that, any
theory which is not falsifiable by an empirical experiment is well
known\footnote{``A theory which is not refutable by any conceivable event is
nonscientific. Irrefutability is not a virtue of a theory (as people often
think) but a vice.'' -- Karl Popper, Conjectures and
Refutations \cite{popper1962conjectures}} to be nonscientific (i.e.
unfalsifiability is a fallacy of a theory), Herley concludes that there is no
scientific theory on cybersecurity; which means that cybersecurity lays in the
realm of pseudo-sciences \cite{Herley2016usenixvideo}.  Herley, e.g.
in \cite{Herley2017justifying}, discusses the implications of a
nonscientific approach to cybersecurity, and highlights the tremendous impact on
all the scientific research and engineering of systems; leading often to
terrorism and wars, and wasting of resources in useless protections or
overspending.  While the criticism is investigated
in \cite{Herley2016unfalsifiability}, no solution is provided nor envisioned.  On the
contrary, the goal of this work is to lay the foundations of a
scientific cybersecurity theory.

We consider the problem raised by Herley not confined to
``computer security'' but to any abstract system (so that our theory may hold
for any sound implementation such as networks, mechanical, cyber, or
cyber-physical system, or even a single computer or a single device such as an
hard-drive).  There is also an apparent inconsistency in
\cite{Herley2016unfalsifiability} that we seek to clarify before following (as
we agree) the scientific path draw by Herley: cybersecurity is defined as an
abstract property in many formal approaches to the investigation of the
security of systems, and the security of the design of a formally verified
protocol is indeed falsifiable (against the security properties verified).  For
example, in the protocol verification community, security is often defined as a
formalization of the high-level properties confidentiality, integrity, and
availability. The problem in such approaches is not the definition of what
cybersecurity is but the generality of the results, since they tackle a specific
step (and not the first) of the engineering process (of systems).  Therefore,
the theories underlying the verification are based on assumptions which
non-evidently apply to a general security theory.  As an example, the so called
Dolev-Yao attacker model\footnote{For the sake of simplicity, the Dolev-Yao
attacker can be considered as an abstraction of an active attacker who controls
the network but cannot break cryptography.} \cite{Dolev1983security} only
applies to specific instances (often called scenarios) and abstraction of the
protocol. This, in turn, creates a false sense of security since requires
non-justifiable assumptions on the abstraction of the system of which security
is verified. More specifically, for the formal security verification of a
system in which two or more agents are communicating, a formalized scenario
needs to be defined by a modeler who chooses (among others): (i) a scope of the
formalization (e.g.  excluding the server that distributes the public key is
often done when verifying the security of authentication protocols), (ii) the
number of sessions (even tough some approaches do reason on an infinite number
of sessions such as \cite{Escobar2007maudenpa}), (iii) honesty/dishonesty of
the peers (e.g.  in the ASLan++ language \cite{Oheimb2010aslan++}), and (iv)
the abstraction of the cryptographic primitives (e.g.  ProVerif vs CryptoVerif
\cite{Blanchet2017symbolic}).  While many tools and theories improves those 
issues, there is no agreement on which should be the definitive
approach (if any). More importantly, some of the choices made by the
modelers/engineers using those formal approaches may completely change the
results of the formal verification of the system (an interesting example on how
difficult it is to even just compare the approaches is given in
\cite{Cremers2009comparing}). 
\begin{figure}[t]
	\centering
	\includegraphics[width=.7\columnwidth]{protocol-example.pdf}
	\caption{Abstraction of an ad-hoc esemplificative protocol execution}
	\label{fig:protocol-example}
\end{figure}
For example, in Figure~\ref{fig:protocol-example}, under the perfect
cryptography assumption\footnote{As defined in \cite{Rocchetto2016cpdy}: ``In
the so called perfect cryptography assumption, the security encryption scheme
is suppose to be perfect, without any exploitable flaw, and so the only way for
the attacker to decrypt a message is by using the proper key. That assumption
is widely accepted in the security protocol community, and most of the formal
reasoning tools for the analysis of security protocols abstract away the
mathematical and implementation details of the encryption scheme
\cite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}''}
and assuming that no violation to any security property is done after message
I), the freedom of choosing the scope
determines that the flaws related to the dishonest impersonation of the Server
may or may not be considered in the verification process.  This choice has
tremendous impact on the focus and findings of the verification of the security
of the protocol.  While this may seem to turn upon minutiae and foreseeable,
this highlights the false sense of security that may derive from a
non-falsifiable theory of system security\footnote{``To the superficial
observer, the analysis of these forms seems to turn upon minutiae. It does in
fact deal with minutiae, but they are of the same order as those dealt with in
microscopic anatomy.'' -- Karl Marx, Capital Volume 1, 1867}.

\paragraph{Structure} We start, in
Section~\ref{sec:literature}, by detailing the problem statement, reporting a
literature review on the main concepts and definitions related to security.  We
formulate a security hypothesis in Section~\ref{sec:hypothesis}; which we use
to propose a theory on system security in Section~\ref{sec:theory}. In
Section~\ref{sec:secra}, we apply our theory for the tool-assisted security
risk assessment of a CPS (with an ad-hoc example, based on the SWaT testbed
\cite{Mathur2016swat}).  This application shows how our theory can be used to
predict all of the possible security weaknesses of a system, allowing the
falsification of our theory.  In fact, if any security weaknesses were to be
found in a system and not predicted by our theory, the theory could be declared
incomplete.  Similarly, if a security weakness would be predicted by our theory
but found to be impossible to have our theory could be declared as wrong.

\section{Literature Review}\label{sec:literature}
\begin{figure*}[t]
	\centering
	\includegraphics[width=.9\textwidth]{safety-security_3.pdf}
	\caption{Overview of keywords related to security and safety}
	\label{fig:safety-security}
\end{figure*}

\fixnote{mr}{add weakness-error in picture and text}
In most of the natural languages the concepts of safety
and security are not syntactically differentiated and both terms (safety and
security) are expressed by the same word, e.g. sicurezza in Italian.  A
semantic distinction between safety and security is correlated to a
belief\footnote{A belief has to be intended as a proposition which is supposed
to be true by the majority of people in our society, without a scientific
underlying theory but based on partial empirical evidences or inductive reasoning based on partial empirical evidences.} that
safety deals with \emph{accidents} (i.e. an unfortunate incident) posed by the
natural environment (e.g. natural events such as wearing of hardware
components) while security deals with \emph{incidents} posed by mankind (e.g.
attackers and bugs).  The fundamental difference between nature and mankind (and,
in turn, between safety and cybersecurity) is believed to be on the different
intents\footnote{``The belief–desire–intention software model (BDI) is a
software model developed for programming intelligent
agents.''\cite{wiki-bdi}. In the BDI model, the intents represents the
deliberative state of an agent which determines the choice of that agent on
what to do.} (accidents are unfortunate while incidents are not) of the causes
that generates a threat; namely, nature is believed not to have malicious
intents (but unfortunate causes-effects) while threats generated by mankind are
malicious\footnote{Of course, logical flaws or bugs may be
introduced by other means (e.g. ignorance) without explicit malicious intents,
but the exploitation of those flaws is considered (for now, and detailed
afterwards in the article) malicious, and then we consider any vulnerability to
be malicious even if due to the lack of skills.}.
This conclusion seems to be against a general formulation of a cybersecurity theory;
how can we define a theory that predicts what a human being will do?
The cybersecurity attacks seems to be related to the creativity of the attacker
and then unpredictable. Currently, our understanding of cybersecurity is
stored into a network of databases of weaknesses (e.g.
CWE\cite{MITRE2020CWEresearch}), vulnerabilities (e.g. CVE\cite{CVE},
NVD\cite{NIST2020NVD}),
and attacks (e.g. CAPEC\cite{MITRE2020CAPEC}, ATTACK\cite{attack}). Systems are
nowadays tested against 
known attacks ignoring the fact that the considering a system secure
if no known attack is possible, defines an unfalsifiable claim.
As an example, if we have a system which is secure against all of the
vulnerabilities in the CVE we just need to wait for a month to have 1000+ more
vulnerabilities to test \cite{newVulns}. If we suppose that one of those new vulnerabilities
affects our system, that should be the falsification of the hypothesis that a
system is secure if no known attacks apply. So, even tough the field 
of security protocols is still lacking of a general cybersecurity theory
the secure-by-design approach they apply can be considered as going
into the opposite direction of the ``implement and test'' approach.

Most of the safety-preserving principles in the field of engineering of
safety-critical cyber-physical systems (such as elevators and aircraft), upon
which safety requirements are defined (e.g. in standards such as the IEC 61508
or 61511\cite{IEC201761511}), are based on empirical tests and measurements
(therefore should be considered hypothesis and not definitions). While
reasoning by induction\footnote{``So, whenever they argue ``Every man is an
animal and Socrates is a man; therefore Socrates is an animal,'' proposing to
deduce from the universal proposition ``every man is an animal'' the particular
proposition ``Socrates therefore is an animal,'' which in fact goes (as we have
mentioned) to establish by way of induction the universal proposition, the
fall into the error of circular reasoning, since they are establishing the
universal proposition inductively by means of each of the particulars and
deducing the particular proposition from the universal syllogistically.''
Sextus Empiricus, Outlines of Pyrrhonism II-195 \cite{Empiricus1990Pyrrhonism}}
based on the empirical observation should be avoided, since it may easily lead
to false beliefs, this approach is often justified by the supposed
impossibility of defining a theory that correctly predicts failures.  A failure
of a wire due to environment (e.g. due to humidity, dust, heat \&c) is defined
from empirical evidences and processes have been standardized to test qualities
of hardware components.  This process completely breaks down when a malicious
environment (i.e. an attacker) is considered instead of the (supposedly honest
and predictable) natural environment. Therefore, the same approach that is in
use for safety, seems not to be applicable for security (e.g. for security
testing).
An overview on the aforementioned aspects of safety and security is depicted in
Figure~\ref{fig:safety-security} and is used as a baseline for a definition of
the terms that structure our current understanding of security. 
\begin{itemize}
	\item \emph{Mankind} ``refers collectively to humans''
\cite{wiki-mankind}, while the concept of \emph{Nature} is
		related ``to the intrinsic characteristics that plants,
		animals, and other features of the world develop of their own
		accord'' (e.g. the physical universe)\cite{wiki-nature}. 
		\begin{itemize}
			\item There exist several terms to refer to an
				\emph{attacker}, i.e. threat agent or threat source,
				considering those terms to be semantically
				equivalent.  From now on, we consider the
				Causality principle to be the \emph{threat
				source}, Nature or Mankind to be the
				\emph{threat agents} and an \emph{attacker} as
				a specific malicious threat agent which materializes a
				threat.
		\end{itemize}
	\item \emph{Vulnerability}\footnote{The term vulnerability is not
		present in the Encyclopedia of Cryptography and Security, while
		it is used in 12 entries (such as in the definition of
		``penetration testing'' \cite{caddy2005pentest})
		highlighting how commonly this word is used without a proper
		supporting semantics.}, as defined in \cite{cnssi20104009}
		(and adopted in \cite{nist2013800-53}), is ``weakness in an
		information system, system security procedures, internal
		controls, or implementation that could be exploited by a threat
		source''. On the one hand, the definition is broad to enclose
		as much causes (that generates a vulnerability) as possible
		On the other hand, the term vulnerability should have a complete and sound
		definition, so that no other causes (e.g.  other sources) but
		the ones in the definition are responsible for a vulnerability
		\footnote{``For this view, that
		\emph{That Which Is Not} exists, can never predominate. You
		must debar your thought from this way of search, nor let
		ordinary experience in its variety force you along this way,
		(namely, that of allowing) the eye, sightless as it is, and the
		ear, full of sound, and the tongue, to rule; but (you must)
		judge by means of the Reason (Logos) the much-contested proof
		which is expounded by me.'' -- Parmenides of Elea, On Nature
		(circa 500 B.C.), fragments B7.1–8.2
		\cite{Hakim2016philosophy}}.
		Furthermore, the term ``threat sources'' used in the definition
		in\cite{cnssi20104009} may be identified with both Nature
		and Mankind, not differentiating between safety and security.
\end{itemize}

As depicted in Figure~\ref{fig:safety-security}, a vulnerability does not
necessarily become a threat for the system, unless exploited ``through a
channel that allows the violation of the security policy
[\ldots]''\cite{cnssi20104009}. For example, a software or procedure that takes
advantage of the vulnerability causing an \emph{attack} to the system may
result in several correlated incidents and threats.  The process of
exploitation of a defect as a vulnerability is reported in
Figure~\ref{fig:safety-security} such that the difference between exploit and failure,
and attack and accident, is to be found just in the maliciousness of the intents
that causes this process (i.e. excluding the intent, the terms are just syntactic transformation from a vulnerability to defect, from
accident to incident).

\begin{itemize}
	\item \emph{Weakness}. The definition given by the MITRE
		in \cite{MITRE2020CWEweakness} of weakness is: ``Software
		weaknesses are errors that can lead to software
		vulnerabilities. A software vulnerability, such as those
		enumerated on the Common Vulnerabilities and Exposures (CVE)
		List, is a mistake in software that can be directly used by a
		hacker to gain access to a system or network''.  The definition
		is circular if we interpret the word ``error'' and ``mistake''
		with the same semantics: a weakness is an error that leads to a
		vulnerability and a vulnerability is a mistake which, in turn,
		is a weakness.  
\end{itemize}

The only difference between a weakness and vulnerability seems to be that one
can consider weakness as a ground term and state that a vulnerability is caused
by a weakness.

\begin{itemize}
	\item \emph{Causality} refers to the causality principle; defined
		in\cite{Spirkin1983Dialectical} as ``Causality is a genetic
		connection of phenomena through which one thing (the cause)
		under certain conditions gives rise to, causes something else
		(the effect). The essence of causality is the generation and
		determination of one phenomenon by another. In this respect,
		causality differs from various other kinds of connection, for
		example, the simple temporal sequence of phenomena, of the
		regularities of accompanying processes''. One can consider the
		causality principle to be formalized by the K Modal Logic.
	\item An \emph{Exploit}\footnote{We note that the term exploit is only
		used as a verb in\cite{ISO2009information}} ``[\ldots]
		(from the English verb to exploit, meaning to use something to
		one’s own advantage) is a piece of software, a chunk of data,
		or a sequence of commands that takes advantage of a bug or
		vulnerability to cause unintended or unanticipated behavior to
		occur on computer software, hardware, or something electronic
		(usually computerized).''\cite{wiki-exploit}.
	\item An \emph{Attack}, as defined by the International Standard
		ISO/IEC 27000 is an ``attempt to destroy, expose, alter,
		disable, steal or gain unauthorized access to or make
		unauthorized use of an asset''; where an \emph{Asset} is
		``anything that has value to the organization''. We note that for
		the purpose of this article, we do not want to focus on a specific
		organization or business to define asset but, in general, on any 
		abstract organization (e.g. a company or a society).
		We do not consider ethical hackers as attacking a system. 
		In fact, we consider the term \emph{hack} as
		non-malicious (as, e.g. in \cite{Stallman2002hacker}).
	\item A \emph{Threat}, as defined in\cite{cnssi20104009}, is ``Any
		circumstance or event with the potential to adversely impact
		organizational operations (including mission, functions, image,
		or reputation), organizational assets, individuals, other
		organizations, or the Nation through an information system via
		unauthorized access, destruction, disclosure, modification of
		information, and/or denial of service''.
	\item \emph{Defect}, ``anything that renders the product not reasonably
		safe''\cite{Robinson2019writing} (i.e. a characteristic of
		an object which hinders its proper usability).
	\item \emph{Failure}, as defined in\cite{Merriam2020failure} as ``a state of
		inability to perform a normal function''. The term is
		structured and detailed in
		\cite{cnssi20104009,iet2017glossary} but relying on an
		abstract notion of failure without a specific definition.
	\item \emph{Hazard}, ``a potential source of
		harm''\cite{iet2017glossary}.
\end{itemize}

\begin{figure}[t]
	\centering
	\includegraphics[width=.7\columnwidth]{causality.pdf}
	\caption{Etiology of cybersecurity}
	\label{fig:causality}
\end{figure}

Our literature review shows that most of the definitions relates insecurity to
dis-honesty (also called maliciousness or adversarial) of an agent (often
called adversary or attacker). This, however, just moves the problem of
defining what cybersecurity is as the problem of defining what an dishonest
agent can or cannot do in a system\footnote{where an agent is any virtual or
physical entity of the system or using the system (e.g. a device, a software,
or a human being) and dishonesty is not necessarily related to malicious
motivation but also to incompetence or lack of skills.} As in the Dolev-Yao
theory, we may correlate ``being dishonest'' to ``not following the intended
behavior/rules''.  In the case of a generic system, a dishonest agent is,
therefore, any agent that doesn't follow the intended behavior (or
functionality or logic) of the system.  Given the generality of the definition,
and its high level of abstraction, we may conclude that the hypothesis seems
evident. For example, a software can be considered an agent of the system and
whenever it has a bug, it can be exploited causing an Incident. However, this
hypothesis has lead to the un-testable conclusion that the dishonest behaviors
of agents cannot be defined in general (e.g. due to the heterogeneity of agents
and systems) and that huge repository of dishonest behaviors should be kept as
definition, such as CAPEC\cite{MITRE2020CAPEC}; or that dishonesty of an agent
with respect to a security protocol should be defined as a number of predefined
actions as in
\cite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}
(to name a few). As depicted in Figure~\ref{fig:causality}, in order to define
a theory on cybersecurity and on the malicious behaviors of agents we noticed
that Attack Patterns (or Attacks) are caused-by the presence of Vulnerabilities
in the system. Those Vulnerabilities are, in turn, cause-by the presence of
Weaknesses in the system.  Weaknesses are Errors in the design or
implementation of a system.  Therefore, a theory on cybersecurity should first
predict the Errors in a system design.

\section{Security Hypothesis -- ABF Theory of Agents}\label{sec:hypothesis}
In order to address the problem raised by Herley, we shall define how to
distinguish between a secure and an insecure system. While most of the
literature have correlated the problem of in-security
to the maliciousness of agents interacting with the system, we show that
security doesn't seem to stem from a malicious nature but, rather, insecurity
raises from the lack of well-defined security requirements for the development
process of a system. We argue that the high number of security vulnerability
reported today are simply the realization of potential system configurations,
which deviate from the nominal behavior because the \emph{intended (or
nominal) behavior} of the system is not precisely defined in the specification
at the very early stages of the engineering process.

As a reference, we consider the following steps of the engineering process
and our objective is to be able to predict the security requirements
instead of axiomatically assume them.
\begin{enumerate}
	\item \emph{System Specification}, where the \emph{functional and physical requirements} are defined
	\item \emph{Architecture Design}, where the specification is structured into \emph{functional and physical architectures}
	\item \emph{Security Risk Assessment}, where the potential \emph{weaknesses} (errors)
		are identified and, consequently, \emph{security requirements} are
		captured.
	%\item \emph{Design Refinement}, where the physical and functional
	%	architectures are extended with \emph{mitigations} for the
	%	weaknesses. Mitigations at this step details the semantics of
	%	the functional blocks (behaviors), the data flow, and communication logic.
	%\item \emph{Verification of Security Requirements}, where the data flow, behaviors, and communication protocols are verified against the security requirements.
\end{enumerate}
Given that, in our investigation, we changed the focus from the attacker to the
potential designs of a system, we start by defining a general framework for the definition
of a system. On top of this framework we will identify system weaknesses as potential
design errors.

\subsection{Systems as Multi-Agent Systems}\label{sec:system}
The ISO/IEC/IEEE 15288:2015 (System Life Cycle Processes) provides a definition
of system as ``A combination of interacting elements organized to achieve one
or more stated purposes.''\autocite{ISO201515288}.  Therefore, a system can be
considered as a single agent where its interacting elements are the
constituents of the agent itself. For the sake of simplicity we first define a
system as an agent and then extend the definition to a ``combination of
interacting'' agents.

There is no agreement between the research communities (e.g.
Multi-Agent-System, Epistemic Logic) on which are the constituent of an agent
as a system. However, the same ideas revolved around for thousands of years.
Some relevant examples for our objective are the following\footnote{
It is interesting to notice that 
in \autocite{Empiricus1990Pyrrhonism}, the author states: ``The
 logical criterion also may be used in three senses -- of the
 agent, or the instrument, or the ``according to what''; the
 agent, for instance, may be a man, the instrument either sense
 perception or intelligence, and the ``according to what'' the
 application of the impression ``according to'' which the man
 proceeds to judge by means of one of the aforesaid
 instruments.''. The author than proceeds showing that
 even in this framework, knowledge still is non-apprehensible.
}.
\begin{itemize}
	\item In \autocite{Hintikka1962knowledge}, Hintikka describes the
		difference between Knowledge and Belief (as epistemological
		concepts), and the whole Doxastic logic defines in details how
		Beliefs can be formalized.
	\item In \autocite{Hintikka1993Information}, Hintikka describes the concept
		of Information and the difference with Knowledge and Belief.
	\item In \autocite{Santaca2016abf}, the authors defines an agent as a
		tuple of Assertions, Beliefs, and Facts.
\end{itemize}
For our argument, as in \autocite{Santaca2016abf}, an agent is composed by its
knowledge, beliefs, and the information or assertions it provides; where
knowledge is defined, as in \autocite{Steup2020epistemology}, as a set of
proposition known by an agent, such that: (i) knowledge requires belief, (ii)
knowledge require truth, (iii) knowledge must be \emph{properly justified}, and
the only objective of Information is to exchange beliefs between agents.  As
defined in \autocite{Hintikka1993Information}, ``Information is specified by
specifying which alternatives concerning the reality it admits and which
alternatives excludes''. This means that if we consider a propositional
variable (which admits the two alternatives True/False) its information is
defined as Believed to be True/False and not Believed to be the opposite.  Due
to the definition of Information and then with its relation to the
probabilistic correlation to truth/reality, we consider information in relation
with an agent's beliefs. Similarly, we consider beliefs to define the actual
behavior of an agent or a system.

The difference between Knowledge and Information as used by Hintikka and Facts
and Assertions as used in \autocite{Santaca2016abf} are not precisely discussed
in \autocite{Santaca2016abf}.  Furthermore, those epistemological concepts are
difficult to define
\autocite{Gettier2012knowledge}\footnote{``\emph{Theaetetus}: [\ldots] He said
that knowledge was true opinion accompanied by reason, but that unreasoning
true opinion was outside of the sphere of knowledge; and matters of which there
is not a rational explanation are unknowable -- yes, that is what he called
them -- and those of which there is are knowable. [\ldots] \emph{Socrates}:
[\ldots] the primary elements of which we and all else are composed admit of no
rational explanation; for each alone by itself can only be named, and no
qualification can be added, neither that it is nor that it is not, for that
would at once be adding to it existence or non-existence, whereas we must add
nothing to it, if we are to speak of that itself alone.  [\ldots]'' Plato --
Theaetetus 201 \autocite{Plato1914Plato}}.  For example, Hintikka in
\autocite{Hintikka1993Information} states that ``A purely logical definition of
information is impossible''. A detailed formalization of those concepts is out
of the scope of this paper. We now define them abstractly, inspired by the
ABF-theory  defined in \autocite{Santaca2016abf} and, in
Section~\ref{sec:theory} we detail them as engineering concepts.
\begin{enumerate}
	\item We consider Information only when the intention of exchanging 
		that Information is from a sender to 
		recipient is defined; and we call it \emph{Assertion}  
	\item Similarly, the portion of Beliefs we consider for system
		engineering is the one that builds (input) or describes
		(output) the behavior of an agent (strategy rules
		\footnote{``The logical structure of information is one of the
		most basic and one of the most basic and one of the simplest
		thing in the wide and wonderful world of logical analysis. This
		point can be put in a deeper perspective. A distinction
		[\ldots] ought to be made [--] between two kinds of rules (or
		principles) in any strategic activity like knowledge seeking.
		On the one hand you have the rules that define the game, e.g.
		how chessmen are moved on a board. The can be called
		\emph{definitory} rules.  They must be distinguished from rules
		[\ldots] that deal with what is better and what is worse in the
		game in question.  Definitory rules do not say anything about
		this subject. Rules which do can be called \emph{strategic
		rules}'' -- Hintikka in \autocite{Hintikka1993Information}}),
		and
	\item We consider a set of axiomatic \emph{Facts} (definitory
		rules\fixnote{mr}{facts are definitory rule, as they don't
		define how a real system is finally implemented (currently) but
		how it should, trough a series of requirements which may not be
		respected, through insecurity, in the implemented system})
		instead of considering the more general epistemic definition of
		Knowledge. Specifically, Facts describes:
		\begin{itemize}
			\item The functional architecture of each agent
			\item The physical/structural (HW/SW) architecture of each
				subsystem of agents and agent within a
				subsystem
		\end{itemize}
\end{enumerate}
We give a graphical representation of (sub-)system and agent in Figure~\ref{fig:system-agent}
\fixnote{mr}{remove ``functional behavior'' in the figure and leave only ``behavior''}.

\begin{figure}[t]
	\centering
	\includegraphics[width=.9\columnwidth]{system-agent.pdf}
	\caption{Example of system and agent structure}
	\label{fig:system-agent}
\end{figure}

\subsection{Mereo-topological Reasoning}
\begin{table*}[t]
\centering
\begin{tabular}{ccclll} 
\rota{\textbf{RCC3}}&\rota{\textbf{RCC5}}&\rota{\textbf{RCC8}}&\textbf{Terminology} & \textbf{Notation} & \textbf{Definition} \\
\hline
&&&Connects with 			& $\mathit{C}(\mathit{X},\mathit{Y})$ 		& $\mathit{X}\subseteq \mathit{Y}$ \\
&&&Disconnected from		& $\neg \mathit{C}(\mathit{X},\mathit{Y})$		& $\mathit{X}\not\subseteq \mathit{Y}$\\
&&&Part of				& $\mathit{P}(\mathit{X},\mathit{Y})$		& $\forall \mathit{Z} ~\mathit{C}(\mathit{Z},\mathit{X}) \rightarrow \mathit{C}(\mathit{Z},\mathit{Y})$\\
&&&Overlaps			& $\mathit{O}(\mathit{X},\mathit{Y})$		& $\exists \mathit{Z} ~\mathit{P}(\mathit{Z},\mathit{X})\wedge \mathit{P}(\mathit{Z},\mathit{Y})$\\
\Tdot&&&  \textbf{Overlaps Not Equal} 	& $\mathit{ONE}(\mathit{X},\mathit{Y})$		& $\mathit{O}(\mathit{X},\mathit{Y}) \land \neg \mathit{EQ}(\mathit{X},\mathit{Y})$ \\
\Tdot&\Tdot&\Tdot& \textbf{Equal to} 		& $\mathit{EQ}(\mathit{X},\mathit{Y})$  		& $\mathit{P}(\mathit{X},\mathit{Y}) \wedge \mathit{P}(\mathit{Y},\mathit{X})$\\
\Tdot&\Tdot&\Tdot& \textbf{DiscRete from} 		& $\mathit{DR}(\mathit{X},\mathit{Y})$		& $\neg \mathit{O}(\mathit{X},\mathit{Y})$\\
&\Tdot&\Tdot&\textbf{Partial-Overlap}	& $\mathit{PO}(\mathit{X},\mathit{Y})$ 		& $\mathit{O}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{X},\mathit{Y}) \wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\ 
&\Tdot&&\textbf{Proper-Part-of} 	& $\mathit{PP}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{X},\mathit{Y})\wedge \neg \mathit{P}(\mathit{Y},\mathit{X})$\\ 
	&\Tdot&&\textbf{Proper-Part-of-\textit{\textbf{i}}nverse} & $\mathit{PPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{P}(\mathit{Y},\mathit{X}) \wedge \neg \mathit{P}(\mathit{X},\mathit{Y})$\\
&&\Tdot&\textbf{Externally Connected} 	& $\mathit{EC}(\mathit{X},\mathit{Y})$ 		& $\mathit{C}(\mathit{X},\mathit{Y}) \wedge \neg\mathit{O}(\mathit{X},\mathit{Y})$\\ 
&&\Tdot&\textbf{Tangential PP} 	& $\mathit{TPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\ 
&&\Tdot&\textbf{Tangential PPi} 	& $\mathit{TPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{TPP}(\mathit{Y},\mathit{X})$\\ 
&&\Tdot&\textbf{Non-Tangential PP} 	& $\mathit{NTPP}(\mathit{X},\mathit{Y})$ 		& $\mathit{PP}(\mathit{X},\mathit{Y})\wedge\neg\exists\mathit{Z}~[\mathit{EC}(\mathit{Z},\mathit{X}),\mathit{EC}(\mathit{Z},\mathit{Y})]$\\ 
&&\Tdot&\textbf{Non-Tangential PPi} 	& $\mathit{NTPPi}(\mathit{X},\mathit{Y})$ 		& $\mathit{NTPP}(\mathit{Y},\mathit{X})$\\ 
\end{tabular}
\caption{RCC3, RCC5, and RCC8 relations between Regions $X$, $Y$ and $Z$ ~\label{tab:rcc358}~\label{tab:rcc}}
\end{table*}

Similarly to\autocite{Santaca2016abf}, we define agents as a meronomy (an
hierarchy of Part-Whole relations) over the constituent that we previously defined
(assertions, beliefs, and facts), based on a standard definition of mereology,
i.e. based on the definition of Parthood relation between \emph{Parts}.  Due to
the necessity of considering different types of Part (as we'll show afterwards)
we extend the mereology to a
mereo-topology\autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
considering the relations in Table~\ref{tab:rcc358}.  For the sake of
readability, we use the term \emph{Region} both to refer to a mereological Part
and to a topological Region.  Our aim is to create a meronomy
instead of the taxonomies such as the one provided
in\autocite{NIST2020NVD,MITRE2020CVE} with the
CVSS\autocite{Mell2007CVSS} scoring system.

A mereotopology, as defined e.g. in\autocite{Rachavelpula2017mereotopology}, is
an ordered mathematical structure where the basic relation between Regions is
the reflexive and symmetric\fixnote{mr}{I guess it must be monotonic as defined
in\autocite{Rachavelpula2017mereotopology} but I don't find it consistently in
other papers.} \emph{Parthood} relation $\subseteq$.  
\begin{definition}{\bf Parthood --}\label{def:parthood}
Given any pair of mereotopological Regions $X$ and $Y$,
	\begin{enumerate}[noitemsep]
		\item Reflexivity: $\forall X.~ (X\subseteq X)$
		\item Symmetry: $\forall X, Y.~ (X\subseteq Y \Rightarrow Y\subseteq X)$
	\end{enumerate}
\end{definition}

The Parthood relation orders a universe of agents $\agentuniverse$ by defining the so called
\emph{Connects with} (see in Table~\ref{tab:rcc358}) relation between Regions.
We use the
Region Connection Calculus (RCC), as defined
in~\cite{bennettLogics,improvingRCC}, to provide an axiomatization of the
mereo-topological concepts. In its broader definition, the RCC theory is composed
by eight axioms, and is known as RCC8. In the text, for brevity, 
we will often focus only on RCC5 (without
loss of generality) by not considering tangential connections between spatial
Regions. In Table~\ref{tab:rcc}, we
summarize the axioms of the Region Connection Calculus (see, e.g., \autocite{Grutter2008rcc}).
We can now define a system over the mereotopology using the RCC calculus, as follows.

\begin{definition}{\bf System State --}\label{def:opsystem}
	A CPS system (or a sub-system) state is defined as the tuple
	$s=\langle\rcc(\factRegion,\beliefRegion),\rcc(\factRegion,\assertionRegion),\rcc(\beliefRegion,\assertionRegion)\rangle$,
	where $\assertionRegion,\beliefRegion,\factRegion$, are regions of
	assertions, beliefs (i.e. the beliefs generated by the behavior), and
	facts expressed as requirements respectively.
\end{definition}

As already presented in \autocite{Santaca2016abf} it follows that, defining
a system (or an operational system) with a fixed number of regions, there exist
an upper-bound to the number of possible configuration of a system, defined by
the possible relations between the different regions.
For completeness, we report in the next paragraph 
the calculation done in \autocite{Santaca2016abf}.

\paragraph{Number of different configurations of a system}
The general formula to calculate the number of different types of agents is
$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
between $n$ different sets, where $r^e$ is the number of permutation of $r$
relations over $e$ elements with repetitions, with $e$ being the number of
$k$-ary combinations of $n$ sets, $\binom{n}{k}$.
In our case, $\binom{n}{k}=3$ since we consider $3$ sets
($\assertionRegion,\beliefRegion,\factRegion$), and all the relations
considered in the RCC are binary.  Hence, using RCC5 (with five different
spatial relations) over three sets, we can theoretically define up to 125
different type of agents. However, only 54 of the 125 (as showed in
\cite{improvingRCC}) combinations are topologically correct with respect to
the definition of the relations of RCC5. Generalizing to all the RCCs:

\begin{itemize}%[nosep]
\item \emph{RCC3} --- theoretical: $3^3=27$,  correct: 15 
\item \emph{RCC5} --- theoretical: $5^3=125$, correct: 54
\item \emph{RCC8} --- theoretical: $8^3=512$, correct: 193
\end{itemize}

Hence, even if considering a different number of sets than the three
$\assertionRegion$, $\beliefRegion$ and $\factRegion$ exponentially affects
the number of theoretical agents, the application of RCC downscales that number
of a factor that ranges from 1.8 to 2.5. In addition, using RCC5 we consider
3.6 times more (different) types of agents than RCC3, but using RCC8 would
allow us to consider 3.5 times more different agents.
In the quantitative evaluation of a single agent, depicted in Figure~\ref{fig:quantitative},
we argue that only 1 configuration represents the nominal (expected) behavior 
of the agent while the other configurations are either impossible to 
implement or diverge from the intended nominal behavior. We note 
that, the numbers reported here do not consider the details of the
engineering process and should be considered a limit of an abstract 
representation of the system.
\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{quantitative.pdf}
	\caption{Cybersecurity risk for a single agent}
	\label{fig:quantitative}
\end{figure}

\subsection{Qualitative Evaluation of Agent Space in $\abf$}\label{sec:agentspace}
\begin{table*}[t]
\centering
\begin{adjustbox}{width=.6\textwidth}
\begin{tabular}{r||c|c|c|c|c} 
& \dr{$\assertionRegion$}{$\beliefRegion$} & 
	\po{$\assertionRegion$}{$\beliefRegion$}& 
	\pp{$\assertionRegion$}{$\beliefRegion$} &
	\ppi{$\assertionRegion$}{$\beliefRegion$} & 
	\eq{$\assertionRegion$}{$\beliefRegion$} \\
\hline
\hline %dr
 \multirow{3}{*}{\dr{$\beliefRegion$}{$\factRegion$}} & 
	\cellcolor{abfred} & %dr
	\cellcolor{abf-rg-1}\dr{$\assertionRegion$}{$\factRegion$} & %po
	\cellcolor{abf-rg-2}\multirow{3}{*}{\dr{$\assertionRegion$}{$\factRegion$}} & %pp
	\cellcolor{abf-rg-3} \dr{$\assertionRegion$}{$\factRegion$}& % ppi
	 \cellcolor{abf-rg-4} \\ % eq
& \cellcolor{abfred}\all{$\assertionRegion$}{$\factRegion$}& %dr
	\cellcolor{abf-rg-1}\po{$\assertionRegion$}{$\factRegion$} & %po
	\cellcolor{abf-rg-2}\dr{$\assertionRegion$}{$\factRegion$} & %pp
	\cellcolor{abf-rg-3}\po{$\assertionRegion$}{$\factRegion$} & %ppi
	\cellcolor{abf-rg-4}\dr{$\assertionRegion$}{$\factRegion$}\\ % e q
 & \cellcolor{abfred}& %dr
	\cellcolor{abf-rg-1}\pp{$\assertionRegion$}{$\factRegion$} & %po
	\cellcolor{abf-rg-2}  & %pp
	\cellcolor{abf-rg-3}\pp{$\assertionRegion$}{$\factRegion$} & %ppi
	\cellcolor{abf-rg-4} \\ %eq
\hline %po
 \multirow{3}{*}{\po{$\beliefRegion$}{$\factRegion$}} &
	\cellcolor{abf-rg-1}\dr{$\assertionRegion$}{$\factRegion$} & %dr
	\cellcolor{abf-rg-2} & %po
	\cellcolor{abf-rg-3}\dr{$\assertionRegion$}{$\factRegion$} & %pp
	\cellcolor{abf-rg-4}\po{$\assertionRegion$}{$\factRegion$} & %ppi
	\cellcolor{abf-rg-5} \\ %eq
 & \cellcolor{abf-rg-1}\po{$\assertionRegion$}{$\factRegion$} & %dr
	\cellcolor{abf-rg-2} \all{$\assertionRegion$}{$\factRegion$} & %po 
	\cellcolor{abf-rg-3}\po{$\assertionRegion$}{$\factRegion$} & %pp
	\cellcolor{abf-rg-4}\ppi{$\assertionRegion$}{$\factRegion$} & %ppi
	\cellcolor{abf-rg-5} \po{$\assertionRegion$}{$\factRegion$}\\%eq
 & \cellcolor{abf-rg-1}\pp{$\assertionRegion$}{$\factRegion$} & %dr
	\cellcolor{abf-rg-2} &  %po
	\cellcolor{abf-rg-3}\pp{$\assertionRegion$}{$\factRegion$} & %pp
	\cellcolor{abf-rg-4}& %ppi
	\cellcolor{abf-rg-5}\\ %eq
\hline %pp
 \multirow{4}{*}{\pp{$\beliefRegion$}{$\factRegion$}} &
	\cellcolor{abf-rg-2}\dr{$\assertionRegion$}{$\factRegion$} & 
	\cellcolor{abf-rg-3}&
	\cellcolor{abf-rg-4} & %pp
	\cellcolor{abf-rg-5}\po{$\assertionRegion$}{$\factRegion$} & %ppi
	\cellcolor{abf-rg-6}\\ %eq
 & \cellcolor{abf-rg-2}\po{$\assertionRegion$}{$\factRegion$} & 
	\cellcolor{abf-rg-3}\po{$\assertionRegion$}{$\factRegion$} & 
	\cellcolor{abf-rg-4}\pp{$\assertionRegion$}{$\factRegion$} & %pp
	\cellcolor{abf-rg-5}\eq{$\assertionRegion$}{$\factRegion$} & %ppi
	\cellcolor{abf-rg-6}\pp{$\assertionRegion$}{$\factRegion$} \\ %eq
 & \cellcolor{abf-rg-2}\pp{$\assertionRegion$}{$\factRegion$} & 
	\cellcolor{abf-rg-3}\pp{$\assertionRegion$}{$\factRegion$} & 
	\cellcolor{abf-rg-4} & %pp
	\cellcolor{abf-rg-5}\pp{$\assertionRegion$}{$\factRegion$} & %ppi
	\cellcolor{abf-rg-6} \\ %eq
 & \cellcolor{abf-rg-2}&
 	\cellcolor{abf-rg-3} & 
	\cellcolor{abf-rg-4}& %pp
	\cellcolor{abf-rg-5}\ppi{$\assertionRegion$}{$\factRegion$} & %ppi
	\cellcolor{abf-rg-6} \\ %eq
\hline %ppi
 \multirow{3}{*}{\ppi{$\beliefRegion$}{$\factRegion$}} &
 	\cellcolor{abf-rg-3}\multirow{3}{*}{} &
 	\cellcolor{abf-rg-4}\dr{$\assertionRegion$}{$\factRegion$} &
 	\cellcolor{abf-rg-5}& %pp
 	\cellcolor{abf-rg-6}& %ppi
 	\cellcolor{abf-rg-7} \\ %eq
& \cellcolor{abf-rg-3}\dr{$\assertionRegion$}{$\factRegion$} &
 	\cellcolor{abf-rg-4}\po{$\assertionRegion$}{$\factRegion$} & 
 	\cellcolor{abf-rg-5}\all{$\assertionRegion$}{$\factRegion$}& %pp
 	\cellcolor{abf-rg-6}\ppi{$\assertionRegion$}{$\factRegion$}& %ppi
 	\cellcolor{abf-rg-7}\ppi{$\assertionRegion$}{$\factRegion$}\\ %eq
& \cellcolor{abf-rg-3} & 
	\cellcolor{abf-rg-4}\ppi{$\assertionRegion$}{$\factRegion$} & 
	\cellcolor{abf-rg-5} & %pp
	\cellcolor{abf-rg-6}& %ppi
	\cellcolor{abf-rg-7}\\ %eq
\hline %eq
	\eq{$\beliefRegion$}{$\factRegion$} & 
	\cellcolor{abf-rg-4}\dr{$\assertionRegion$}{$\factRegion$} & 
	\cellcolor{abf-rg-5} \po{$\assertionRegion$}{$\factRegion$} & 
	\cellcolor{abf-rg-6}\pp{$\assertionRegion$}{$\factRegion$} & %pp
	\cellcolor{abf-rg-7} \ppi{$\assertionRegion$}{$\factRegion$} & %ppi 
	\cellcolor{abfgreen} \eq{$\assertionRegion$}{$\factRegion$}  %eq
\end{tabular}
\end{adjustbox}
\caption{RCC5 composition table over 3 regions. The results show that there
exist 54 possible relations and the coloring anticipates the ideal risk matrix
(green the secure state with low risk, red the high risk state, and a gradient
of medium risk states). 
\all{$\assertionRegion$}{$\factRegion$} = \{\dr{$\assertionRegion$}{$\factRegion$}, \po{$\assertionRegion$}{$\factRegion$}, \pp{$\assertionRegion$}{$\factRegion$}, \ppi{$\assertionRegion$}{$\factRegion$}, \eq{$\assertionRegion$}{$\factRegion$}\}
\label{tab:5com}}
\end{table*}
While a quantitative analysis reveals how many possible configurations of an
agent (i.e. a system) exist w.r.t. the $\abf$ theory (i.e. $54/125$ in RCC5), a
qualitative analysis of the different configurations describes the
configurations allowed by the $\abf$ theory, and how those configurations can
be categorized.  In Table~\ref{tab:5com}, we provide the generic composition
table of RCC5 over 3 regions instantiated over $\abf$, which shows the whole
state space for a single agent. The color coding of the table represents the
security risk related to a generic agent.

As depicted in Figure~\ref{fig:soundness}, the relation between Facts, and
Assertions and Behavior defines the soundness of the design (admissible
configurations of Assertions, Behaviors and, in turn, Beliefs) w.r.t. the
specification (what the specification mandates, such as by the nature of the
physical/functional architectural models). We categorize agents by first
analyzing the relations between each pair of Regions defining an agent (i.e.
$\abf$), and then we categorize the different agents as tuple of the three
Regions.  For the sake of simplicity, soundness is opposed to non-soundness in
the following, however, with the RCC one should consider different ``degrees''
of non-soundness. For example, in RCC5, if we consider EQ between two Regions
as representing soundness, DR over the same Regions represents non-soundness;
while PP, PO, PPi represents the different degrees of non-soundness.  A similar
argument can be done for completeness.  \fix{mr}{it's missing the relation with
equivalence. missing parallelism with knowledge-belief/assertion (see slide
v-research-demo)}
\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{soundness.pdf}
	\caption{Example relation between Facts, and Assertions and Beliefs}
	\label{fig:soundness}
\end{figure}

\paragraph{\Rcc{$\assertionRegion$}{$\beliefRegion$} -- Collaboration} In order
to reason on the relation between assertions and behavior we first need to
consider that, by definition, assertions are defined as transfer of information
between two agents, e.g., $a$ and $b$.  Therefore, as depicted in
Figure~\ref{fig:system-agent}, an agent has two main categories of assertions,
input and output assertions.  Given an agent $a$ and a collection of asserted
predicates $\Phi$, the Input assertions are those received by $a$ from an agent
$s$ acting as a sender, $\rassert{s}{a}{\Phi}$; similarly, output assertions
are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. We shall consider
two pairs of regions\footnote{``I am not asserting, as Lotze did, that a
relation between $X$ and $Y$ consists of a quality in $X$ and a quality in $Y$
-- a view which I regard as quite indefensible. -- I assert that a relation $Z$
between $X$ and $Y$ involves the existence in $X$ of the quality ``having the
relation $Z$ to $Y$'' so that a difference of relations always involves a
difference in quality, and a change of relations always involves a change of
quality.'' --  Ellis J. McTaggart, The Unreality of
Time\autocite{Mctaggart1908unreality}}: 
\begin{itemize}
	\item \Rcc{$\assertionRegion_{s\rightarrow a}$}{$\beliefRegion$},
		where the relation between Input-assertions and behavior
		describes the soundness of the execution of the functional
		architecture w.r.t. input elicitation. With more details, the
		ideal specification of the functional architecture, along with
		the expected inputs, defines the functional behavior of an
		ideal system.  If all the inputs (Assertions) are correctly
		handled in the functional specification (behavior) the
		specification is complete.

	\item \Rcc{$\assertionRegion_{a\rightarrow r}$}{$\beliefRegion$},
		where the relation between behavior and outputs describes the
		completeness of the behavior defined in the specification
		w.r.t. the input elicitation.  With more details, if all the
		outputs (assertions) of the functional architecture can be
		produced, the functional architecture is complete.
\end{itemize}

\paragraph{\Rcc{$\assertionRegion$}{$\factRegion$} and
\Rcc{$\beliefRegion$}{$\factRegion$} -- Honesty and Competence} While
Assertions and Beliefs generated by the Behavior determines how a system should
work, and the relation between the two defines the quality (i.e. the
completeness w.r.t the specification) of the agent, the relation of Assertions
and Behavior with Facts determines the quality with respect to the nominal
(specified) system.  Given that Facts defines what must be true in the system,
the relation of Assertions an Facts determines the degree of quality between
the information circulating in a system (or within an agent) and
reality\fixnote{mr}{assertions are reality while facts are dogmatic assertions
at specification level, i.e. requirements}.  Since the transfer of information
through Assertions generates Beliefs, a dishonest agent may circulate false
information, generating false Beliefs.  We note that it is often implied that
the intention behind circulating false information discerns a dishonest and an
incompetent agent, however, we consider honesty related to sharing
truths\footnote{``[\ldots] truth exists only in the good man, but the true in
the bad man as well; for it is possible for the bad man to utter something
true.'' Sextus Empiricus, Outlines of Pyrrhonism,
II-83\autocite{Empiricus1990Pyrrhonism}}.  The relation between Beliefs and
Facts determines the competence (on the subjects defined by the Facts) of an
agent (i.e. the more competent an agent is, the more likely a Belief of that
agent is true).

\subsubsection{$\abf$ Security Enumeration (SE)} The following security
requirement for a CPS specification can be summarized:
\begin{enumerate}
	\item[SE-$1$] Proper interaction between correctly-behaving agents (in
		contrast with the ``Improper Interaction Between Multiple
		Correctly-Behaving Entities'' defined by the CWE--$435$ as one
		of the top ``view'' of the ``research concepts'' in
		\autocite{MITRE2020CWEresearch}) is defined as
		$\eq{\assertionRegion_a}{\beliefRegion_a}$ for an agent $a$
		while can be detailed as follows when multiple agents are
		considered.
	\begin{enumerate}
		\item[SE-$1.1$] The equality relation
			$\eq{\assertionRegion_{s\rightarrow
			a}}{\beliefRegion_a}$ describes the intended secure
			behavior as: the beliefs generated by the behavior of
			the functional architecture shall be complete w.r.t.
			the specified inputs of the agent. Therefore, \emph{the
			assertions received by an agent or a system shall be
			compliant with the expected inputs of the functional
			architecture}. For example, the inputs of the user of a
			SW must be sanitized to exclude deviations w.r.t. the
			expected inputs of the functions implemented in the SW.
			Another example is the type checking between allowed
			inputs and expected inputs.
		\item[SE-$1.2$] Similarly, the equality relation
			$\eq{\assertionRegion_{a\rightarrow
			r}}{\beliefRegion_a}$ defines that the outputs of an
			agent $a$ shall be the outputs of the functional
			architecture.
	\end{enumerate}
	\item[SE-$2$]Sufficient Control Flow Management (in contrast with
		the ``Insufficient Control Flow Management'' defined by MITRE
		in the CWE--$691$ as one of the top ``view'' of the ``research
		concepts'' in\autocite{MITRE2020CWEresearch}) is defined as
		$\eq{\assertionRegion}{\factRegion}$.
	\item[SE-$3$] Correct Calculation (in contrast with the ``Improper
		Calculation'' defined by MITRE in the CWE--$682$ as one of the
		top ``view'' of the ``research concepts''
		in\autocite{MITRE2020CWEresearch}) is defined as
		$\eq{\beliefRegion}{\factRegion}$.
\end{enumerate}

We are now in the position to define what a secure system is (with respect to
the $\abf$-theory) and, based on that definition, what the security risk is and
how to quantify it in a risk matrix.
\begin{definition}{\bf Security of a System or an Agent --}\label{def:security}
	A secure system is a system where SE-$1$, SE-$2$, and SE-$3$ holds for
	each agent composing the system.
\end{definition}
The ISO 31000 consider risk as the ``effect of uncertainty on objectives'' and
the refers both of positive and negative consequences of uncertainty.
Accordingly, we consider risk as follows. 
\begin{definition}{\bf Risk --}
The whole space of potential designs of a specification with respect to the
	$\abf$-theory.
\end{definition}
The definition of Risk leads to the risk matrix in
Figure~\ref{fig:quantitative}, defined as follows.
\begin{definition}{\bf Risk Matrix --}
	The risk matrix is a function between the three relations
	$s=\langle\rcc(\factRegion,\beliefRegion),\rcc(\factRegion,\assertionRegion),\rcc(\beliefRegion,\assertionRegion)\rangle$,
	where the maximum risk is defined by the $DR$ relation between the
	three groups of Regions, and the minimum risk by the $EQ$ relation over
	the same Regions. In between the two extremes, the granularity of
	possible intermediate configuration is defined by the calculus used
	(RCC5 in our case).
\end{definition}
We note that often the risk matrix is defined by the relation between
likelihood and impact which we discuss in the next section.

\section{Prediction of Security Weaknesses}\label{sec:theory}
Several standards mandates a secure-by-design approach in which cybersecurity
shall be considered at the very early stages of the design process.  For
example,
\begin{enumerate}[noitemsep]
	\item DO-326A -- ``Airworthiness Security Process Specification''
		requires a cybersecurity risk assessment of the design and is
		the ``are the only Acceptable Means of Compliance (AMC) by FAA
		\& EASA for aviation cyber-security airworthiness certification,
		as of 2019'' as pointed out by SAE in \autocite{SAE2019DO326A}.
	\item NIST 800-82 \autocite{Stouffer2011guide} -- ``guide to Industrial
		Control System (ICS) Security''
	\item J3061:2016-1 \autocite{SAE2016J3061} -- ``Cybersecurity Guidebook
		for Cyber-Physical Vehicle Systems'' defines ``set of
		high-level guiding principles for Cybersecurity as it relates
		to cyber-physical vehicle systems'' and states that
		``incorporate Cybersecurity into cyber-physical vehicle systems
		from concept phase through production, operation, service, and
		decommissioning''
\end{enumerate}
However, standards do not describe in detail how to perform a security risk 
assessment and only vaguely define the overall objective. The overall
objective of this first step of the cybersecurity engineering process
is to provide an understanding of the potential security risks.
Roughly speaking, a security risk assessment (a.g. \autocite{coras,secram}) 
methodology starts from the (i) identification of assets, (ii) determines
the threat scenarios (correlating vulnerabilities, threats and incidents) e.g. 
threat diagrams in the CORAS approach, (iii) calculate the risk as a function
between impact and likelihood of threat scenarios, and (iv) finally provides
treatments (mandating a partial re-design) and residual risks.
All the methodologies and tools we reviewed (Threatmodeler, CORAS, SECRAM, ...)
relies on the expertise of the person who performs the risk assessment for
the identification of threats and for the quantitative estimation of risks.
As Herley puts it \autocite{cite}, ``we don't have a test for security and
we don't have a metric for security'' while standards requires a security risk assessment
which, in turn, mandates a metric to evaluate the security risk.

So far, we have reasoned on systems in the abstract, considering
generic MAS; mapping Epistemological concepts
to MAS. We now map MAS to CPS and show how this mapping allows us
to predict security weaknesses in a system.

\subsection{From Multi-Agent to Cyber-Physical Systems}
\begin{figure*}[t]
	\centering
	\includegraphics[width=.9\textwidth]{mashyp.pdf}
	\caption{asd}
	\label{fig:mashyp}
\end{figure*}

As depicted in Figure~\ref{fig:mashyp}, we relate MAS and CPS as follows.
\begin{itemize}
	\item We consider a System as an hierarchy of agents. So, we map agents
		to \emph{systems, sub-systems, or devices}, depending on the
		granularity of the design. For example, a modeler can model a
		specific device, as in Figure~\ref{fig:mashyp}. In this case,
		the system is not decomposed into sub-systems or devices and
		the device is considered as an agent.
	\item Agents reason over Beliefs (i.e. transform Beliefs into other
		Beliefs) and for each \emph{component of a CPS} (system, sub-system,
		or device) is composed by a \emph{functional architecture} that
		transforms input-Beliefs into output-Beliefs.
	\item Components of a CPS have \emph{ports} to exchange information
		with the outer environment (which may be a sub-system),
		similarly to Agents. The transfer of information in a CPS is
		defined by \emph{channels}.
	\item The concept of Knowledge is related to the \emph{requirements}
		that, in a factorial way, describe how the CPS shall behave and
		its \emph{physical architecture}.

\end{itemize}

\paragraph{Input and Output Ports}
\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{IOports.pdf}
	\caption{Input and Output Port of an agent}
	\label{fig:ioports}
\end{figure}
Since the $\abf$-theory is a theory of agents, we could consider ports as
agents that allows the exchange of information between a channel and another
agent.  However, we considered a port as a special type of agent that to avoid
an \emph{infinite regress} in which a port needs a port to transfer information
between the outside of the port to the inside of itself.

\begin{definition}{\bf Input or Output Port --}\label{def:port} 
	A port forwards information from the outside of an agent's boundary to
	the inside (input-port) or vice versa (output-port).  There exist two
	types of ports with the following behavior: an input-port changes
	transforms Assertions from a sender $s$ ($\rassert{s}{a}$) to Beliefs
	of an agent $a$ ($\beliefRegion_a$), while an output-port from belief
	to assertion.
\end{definition}
The \emph{quality of a port} is determined by the $\rcc$ relation between the
assertions received or sent and the belief, i.e.
$\Rcc{\assertionRegion_s}{\beliefRegion_a}$ for the input-port or
$\Rcc{\assertionRegion_r}{\beliefRegion_a}$. A port is, in fact, syntactic sugar to express the 
relation between Assertions and Beliefs. We only define secure input port
but the definition of a secure output port is symmetrical.

\begin{definition}{\bf Secure Input Port --}\label{def:secport}
	A secure input port ($s\secureport a$) always allow information as
	incoming assertions to flow from a sender $s$ to the behavior of the
	recipient $a$ (agent).  
	%If $p_I$ is a secure input port,
	%$\world\models\varphi$, $\varphi\in\rassert{s}{a}$, and $a$ has the
	%only input port $p_I$, $\world'\models\beliefRegion_a$ and
	%$\varphi\in\beliefRegion_a$ for any $\world\modalrelation\world'$.
\end{definition}

\paragraph{Port Weaknesses}
From our definition of ports, it follows that there exist only the following \emph{six}
types of weaknesses, generating six types of insecure port in RCC5 (the
notation is reported for input-ports, i.e. on the lef-hand side of the arrow):
\begin{enumerate}[start=1, label={W\arabic*)}]
	\item \emph{replace port} ($s\dropport a$), where assertions reaches
		the port but do not pass the boundary of the agent (i.e. do not
		become belief of the agent)
	\item \emph{drop port} ($s\dropport a$), where assertions reaches the
		port but do not pass the boundary of the agent (i.e. do not
		become belief of the agent)
	\item \emph{insertion port} ($s\insertport a$), where some new
		information is believed by $a$ as incoming from the port but
		$s$ didn't send it
	\item \emph{injection port} ($s\injectionport a$), where information
		coming from $s$ is substituted with new information which
		becomes believed by $a$
	\item \emph{selective port}, where some information passes the port and
		part is either:
	\begin{enumerate}[start=1, label={W4.\arabic*)}]
		\item drop ($s\selectivedropport a$), 
		\item drop+insert ($s\selectiveinsertport a$).
	\end{enumerate}
\end{enumerate}

\fixnote{mr}{this proof needs to be restructured as in slides v-research-demo}
\begin{proof}
An input port is, in the $\abf$-theory, defined secure as long as the relation
	between the two regions of input assertions $\assertionRegion$ and
	output beliefs $\beliefRegion$ are equal, i.e.
	$\eq{\assertionRegion}{\beliefRegion}$. Therefore, any other relation
	should result in a weakness (related to an insecurity flaw) of that
	input port.  Using RCC5, there exist exactly other $4$ different type
	of relations, one of which is the discrete-from (DR) relation, i.e.
	$\dr{\assertionRegion}{\beliefRegion}$. When two regions are related by
	the DR relations, they have no subregion in common. Lets
	define a function weight $|X|$ such as, for any region $X$, it
	represents the smallest possible cardinality of a (mereo)topological base for
	$X$; where a base is a collection of regions in a (mereo)topology such that
	every open region can be written as union of elements of that base. 
	We distinguish between Regions that contain Information and Regions that
	don't by writing the latter as $\emptyset$.
	\begin{enumerate}
		\item if $\eq{\assertionRegion}{\beliefRegion})$ then either
			$\assertionRegion=\beliefRegion=\emptyset$ (no
			communication) or
			$\assertionRegion=\beliefRegion\neq\emptyset$ (forward
			communication)
		\item if $\dr{\assertionRegion}{\beliefRegion})$ then
			$\assertionRegion=\emptyset\oplus\beliefRegion=\emptyset$
			(we call \emph{insert} the former and \emph{full drop}
			the latter case), or
			$\assertionRegion\neq\emptyset\wedge\beliefRegion\neq\emptyset\wedge\assertionRegion\neq\beliefRegion$
			which we call \emph{replace} (i.e. drop and insert)
		\item if $\pp{\assertionRegion}{\beliefRegion})$ then
			$\beliefRegion$ contains and extend $\assertionRegion$
			which we call \emph{injection}
		\item if $\ppi{\assertionRegion}{\beliefRegion})$ then
			$\assertionRegion$ contains and extend $\beliefRegion$
			which we call \emph{selective drop}
		\item if $\po{\assertionRegion}{\beliefRegion})$ then a part of
			the $\assertionRegion$ is contained in the
			$\beliefRegion$ which is a combination of
			\emph{selective drop and generation}
	\end{enumerate}
\end{proof}

\paragraph{Communication Channels}
\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{channel.pdf}
	\caption{Communication over a Mono-directional Channel}
	\label{fig:channel}
\end{figure}
In this work, we only consider
mono-directional channels and communication but the extension to bi-directional
channel can be considered as the union of two unidirectional channels. A
mono-directional channel is defined by the assertions sent or received (over
the channel).

We start by considering the difference between a (communication)
mono-directional channel (channel from now on) and an agent, as we did for the
ports, since the $\abf$-theory is a theory of agents.  In fact, if a channel
were considered an agent (channel-agent) then the question would be how an
agent would transfer its Assertions to the channel-agent. If the channel
between the agent and the channel-agent is again an agent, we would generate an
\emph{infinite regress}. Therefore, we do allow channel-agents but we assume a
finite depth (of detail) for a channel, where there exists a bottom-channel
which is not an agent. For now, we do not constrain a channel-agent in any way
so there is no difference between a channel agent and agent. Therefore,
we consider channels to be bottom-channels,
defined as agents with the pre-defined behavior (i.e. defined in a axiomatic
way) of forwarding any input-assertion as output-assertion, without modifying it\footnote{Nothing
prevents us from introducing additional constraints to the channel as storing
Assertions that are transferred over the channel, or filter out some
input-assertions.}.

\begin{definition}{\bf Secure Mono-directional Channel (bottom-channel) --}\label{def:monochannel}
	A mono-directional channel between the two agents ($s \rightarrow r$) is defined as 
	an agent which behavior is (dogmatically) defined as: to forward any Assertion received from $s$ over an input-port, to  
	the output-port where $r$ is listening to.
\end{definition}
The \emph{quality of a mono-directional} channel is defined as the $\rcc$
relation between the Assertions made by the sender and the ones received by the
receiver, i.e. $\Rcc{\assertionRegion_s}{\assertionRegion_r}$.

\paragraph{Channel Weaknesses}
Given that a mono-directional bottom-channel is assumed to be perfectly
forwarding any Assertion from its input-port to its output-port, there is no
insecure behavior but only the combination of the weaknesses of the input and
output port; therefore there exists $(7^2)-2=47$ theoretical configurations
($7^2$ because there are $6$ insecure types of port, plus $1$ secure type, on
both input and output side; and we exclude the configuration with $2$ secure
types as input and output, $-2$); where only 43 are possible. For the sake of
readability, we report 6 examples in the following but the proof by exhaustion
over all the possible cases is straightforward.

\begin{enumerate}[start=6, label={W\arabic*)}]
	\item \emph{secure output port} and \emph{input drop port} ($s\sdcom r$), 
	\item \emph{secure output port} and \emph{input insertion port} ($s\sicom r$), 
	\item \emph{output drop port} and \emph{input drop port} ($s\ddcom r$)
	\item \emph{output drop port} and \emph{input insertion port} ($s\dicom r$)
	\item \emph{output drop port} and \emph{input secure port} ($s\dscom r$)
	\item \emph{output injection port} and \emph{input secure port} ($s\jscom r$)
\end{enumerate}

\paragraph{Security Weaknesses} 
It is important to note that all the results of the application of the
$\abf$-theory to channels (the analysis of the relation
$\Rcc{\assertionRegion_o}{\assertionRegion_i}$) lead to the same results of the
analysis of a pair of an input and output port (i.e.
$\Rcc{\beliefRegion_1}{\assertionRegion_2}$ and
$\Rcc{\assertionRegion_2}{\beliefRegion_3}$).  The same result can be obtained
by analyzing the relation between the outputs of a functional block and the
inputs of another functional block, where functional blocks are constituents of
the functional architecture as described afterwards.  As depicted in
Figure~\ref{fig:atom}, so far, we have considered Information generated by a
Port $P_I$ and then sent through a channel $C$ to another (recipient) port
$P_O$. In this scenario where Ports and then Channels are atomic (otherwise
raising infinite regress), we can only  consider the relations between Ports
and Channel; considering both input-port to channel and channel to output-port.
In fact, the Weaknesses of a channel are defined in terms of the Weaknesses of
Ports. We now introduce the concept of functional architecture and its
decomposition into functional blocks.  Io order to define a functional block
without encountering a infinitely recursive definition, we must reach the same
conclusions as for the Channel. Therefore, describing the information as
flowing over a channel or in a functional block is purely syntactic sugar.

We can summarize these results by saying that the relations between Assertions
and Beliefs, Output Assertions of an Agent and Input Assertions of another
Agent, or Output Beliefs of a Functional Block and Input Beliefs of another
Block can \emph{only} be affected by the following weaknesses: replace, drop,
injection, insertion, selective drop, and selective drop + insertion.  We can,
then, deduce the following security properties to mitigate security weaknesses
of a port or a channel (between ports, functional blocks, or both).
\begin{itemize}
	\item \emph{Order-preserving} -- it shall be known if Information is \emph{replaced}
	\item \emph{Availability} -- it shall be known if Information is \emph{dropped} or \emph{selectively dropped}
	\item \emph{Integrity} -- it shall be known if Information is \emph{injected}
	\item \emph{Authentication} -- it shall be known if Information is \emph{inserted}
\end{itemize}
%We can also generalize the weaknesses related to the data flow (i.e. the flow
%of Information as Assertions or Beliefs) as the relation between output-input
%Beliefs between functional blocks, output-input Assertions of a channel,
%inputs-outputs of a port, as follows.
%\begin{itemize}
%	\item \emph{equal}, expected nominal behavior
%	\item \emph{discrete-from}, drops all the inputs and inserts new malicious data
%	\item \emph{proper part}, selectively drops inputs
%	\item \emph{proper part inverse}, forwards all the inputs but crafts and
%	      inserts new malicious data
%	\item \emph{partial overlap}, selectively drops inputs and inserts new data.
%\end{itemize}

\begin{figure}[t]
	\centering
	\includegraphics[width=.9\columnwidth]{engineering_relations.pdf}
	\caption{Relations between Ports and Channel}
	\label{fig:atom}
\end{figure}

\paragraph{Functional Architecture}
A Functional Architecture takes Information as input-Beliefs and transforms the
Information into output-Beliefs. Those transformations occurs within the
Functional Architecture, where Functional Blocks transforms Beliefs into other
Beliefs. Similarly to channels, we could consider a Functional Block as a
Functional Architecture occurring in an infinite regress. Therefore, we
consider Functional Blocks as executing an abstract undefined behavior, of
which we only observe the inputs and the resulting outputs (Beliefs).

\begin{definition}{\bf Functional Block and Architecture --}\label{def:funblock}
	A functional block of an agent 
	takes a region of input-Beliefs and outputs a region of output-Beliefs. 
	A functional architecture is an interconnected system of functional blocks.
\end{definition}
It is evident that the quality of a functional block cannot be determined
by the difference between its inputs and outputs (as we did for
ports and channels). The behavior of a functional block
cannot be determined in general; since any functional block will have 
its own purpose based on functional requirements. 
Therefore, while a port semantics is determined by the relation 
between Assertions and Beliefs, the semantics of a Functional Block 
is determined by the relation between Facts/Requirements and I/O Beliefs.

In other words, a functional block is a generic agent with no pre-defined general
behavior (while ports and channels have a pre-defined behavior).

\begin{itemize}
	\item $\po{\beliefRegion}{\factRegion}$ the component has a Byzantine behavior where occasionally outputs the expected
	        output given the correct inputs. Not all the inputs are handled
	        properly, nor all the expected outputs are always generated
	        when correct inputs are given.
	\item $\pp{\beliefRegion}{\factRegion}$ part of the expected outputs are not
	        generated in response to the correct
	        inputs
	\item $\ppi{\beliefRegion}{\factRegion}$ the components
	        correctly performs the expected behavior when the correct
	        inputs are provided but is subject to input
	        injections
	\item $\dr{\beliefRegion}{\factRegion}$ the component
		never performs the expected behavior (e.g. physical
		damage)
\end{itemize}

\paragraph{Requirements as Facts}
During the specification phase, for any agent, channel, port, functional block
and architecture, there may exist a factorial requirement (Fact) predicating over them.
In other words, any requirement is defined as a Fact since they must be true in
any design or implementation. As we stated in Section~\ref{sec:theory} and depicted in
Figure~\ref{fig:soundness}, Facts are strategic rules that defines how
the system shall behave (by specification), while reality may be shown to
be insecure.

\begin{figure}[t]
	\centering
	\includegraphics[width=0.4\columnwidth]{sum.pdf}
	\caption{Example of a Functional Block}
	\label{fig:sum}
\end{figure}

As an example, consider a functional block that performs the summation of two
inputs, as in Figure~\ref{fig:sum}, defined by the requirement $r := b_3 = b_1
+ b_2$. The possible relations between the behavior of the functional block and
the requirements (i.e. $\Rcc{\beliefRegion}{\factRegion}$ is determined by
the relations between the I/O Beliefs $sum(b_1,b_2)=b_3$ and the requirement
$sum(b_1,b_2)=b_1+b_2$, as follows.
\begin{itemize}
	\item $\eq{\beliefRegion}{\factRegion}=\eq{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, 
		where $\beliefRegion_3$ represents the Region of all the
		outputs of the functional block $sum$ while the Region
		$\beliefRegion_1+\beliefRegion_2$ represents the expected
		outputs of an ideal implementation of the requirement $r$.
		Therefore, the functional block correctly implements the
		requirements.
	\item
		$\dr{\beliefRegion}{\factRegion}=\dr{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$,
		the function block never implements the requirements.
	\item
		$\pp{\beliefRegion}{\factRegion}=\pp{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, there exist some inputs for which
		the output is incorrect.
	\item $\ppi{\beliefRegion}{\factRegion}=\ppi{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$, there exists some outputs that are not the result of the summation of the inputs but given the correct inputs the function always outputs correctly.
	\item
		$\po{\beliefRegion}{\factRegion}=\po{\beliefRegion_3}{\beliefRegion_1+\beliefRegion_2}$,
		the component has a Byzantine behavior where occasionally
		outputs the expected output given the correct inputs. Not all
		the inputs are handled properly, nor all the expected outputs
		are always generated when correct inputs are given.
\end{itemize}

\subsection{Security and Insecurity of a System}
\begin{hypothesis}{\bf System Security Design --}\label{hyp:security}
	A system security design is given by a precise system 
	specification over the physical and functional architectures, that
	uniquely, i.e.  tested against the range of design possible in the
	$\abftheory$-theory, defines the design built on top of those
	requirements.
\end{hypothesis}

\begin{hypothesis}{\bf System Insecurity Design --}\label{hyp:insecurity}
	If, given a system specification as a collection of
	requirements on the system, there exist
	a non-unique design with respect to those requirements, the number of
	equivalent designs that fulfills the requirements quantitatively
	defines the magnitude of insecurity of a system design.
\end{hypothesis}

Based on these hypothesis, security ($S$) can be expressed as the following equation.
\begin{displaymath}
	S=1-(\prod_{a\in Ag} r^{\binom{n}{k}}-\delta_T)
\end{displaymath}
where $r^{\binom{n}{k}}$ represents the number of possible configurations (permutations) 
of an agent based on the number $r$ of relations (e.g. EQ, DR, PO) with arity $k$ between
$n$ different sets in the $\abftheory$ theory. $\delta_T$ represents the 
number of configurations which are not satisfiable with respect to the logical theory
defining the 
algebraic structure (topology) and constraints of the calculus (RCC). $Ag$ is the 
set of all the agents in the system under evaluation.
The constant 1 represents the nominal configuration of the system and then the
only secure configuration. The product represents all the unintended, insecure, configurations.

\subsection{Security Risk Assessment}\label{sec:secra}
In order to test our theory we implemented a tool-chain (open-source under the
AGPLv3 license and available at \autocite{cybersecurityGitHubRepo}) for the
identification of weaknesses and the precise calculation of potential insecure
configurations. The engineering of the $\abftheory$ is summarized in the UML
Class diagram in Figure~\ref{fig:secraclassdiagram} (in appendix).

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{v-secra.pdf}
	\caption{Security Risk Assessment Tool}
	\label{fig:secra}
\end{figure}
As depicted in Figure~\ref{fig:secra}, the security risk assessment process
starts with the definition of the use cases and architectural requirements; the
specification of a system.  In our process, the specification is manually
translated into a UML design where:
\begin{itemize}
	\item a \emph{Deployment Diagram} describes the \emph{Physical
		Architecture}. Each agent is defined as a Node with (physical)
		Ports, and Agent's ports are connected via Information Flow
		connectors, representing the physical channel.
	\item a \emph{Functional Architecture} is linked to each agent in the
		deployment diagram and is defined by an \emph{Object Diagram}.
		The Object Diagram is composed by Instances of Functional
		Blocks, connected via Information Flow connectors.
	\item the connection between the two diagrams is implemented by
		``sockets'', functional blocks connected to a 
		physical port.
\end{itemize}

The tool generates a graph-like internal structure which represents
the specification (called ABF graph). The graph is translated into 
a logical formula that represents the specification in the $\abftheory$-theory and,
along with the axiomatization of the RCC5 calculus, is given as input to the Z3 SMT solver.
The solver is used to identify all possible configurations of the system and, in turn,
to identify all potential weaknesses. 
The internal structures of the tool
can be viewed as PDF and the results are reported into an Excel file.
The Excel file also report the total number of configurations as indicating
the security risk associated to the specification. A user can change the status
of each weaknesses in the Excel file, from the initial (default) status (open)
to ``mitigated''. As a consequence, the risk is re-calculated on-the-fly, i.e.
without the need of running the tool again, based on annotations and formulas 
in the Excel file.

We note that, in this approach, security requirements are not imposed by the
specification but are automatically extracted by our tool as potential weaknesses.

\paragraph{Case Studies}
We report the results of the evaluation of a water level reader (sensor ad-hoc example). 
\begin{figure}
	\centering
	\includegraphics[width=.7\columnwidth]{eng_cs1.png}
	\caption{Water Level Reader Deployment Diagram}
	\label{fig:eng_cs1}
\end{figure}
As in Figure~\ref{fig:eng_cs1}, we defined 2 agents: sensorInTank and sensorBoard,
which represent the physical reader that needs to be placed in a tank of water, and 
the board that interprets the readings and outputs them as digital signals.
The two components are connected by a wire.
\begin{figure*}
	\centering
	\includegraphics[width=.7\textwidth]{internal_cs1.png}
	\caption{Sensor Board Object Diagram}
	\label{fig:int_cs1}
\end{figure*}
In Figure~\ref{fig:int_cs1} we report the functional architecture that
receives the incoming communications from the sensor in the tank, and 
communicates them encrypted.
\begin{figure*}
	\centering
	\includegraphics[width=\textwidth]{results_excel.pdf}
	\caption{Partial View of the results in the Excel file}
	\label{fig:results}
\end{figure*}
The results, summarized in Figure~\ref{fig:results} reports more than 16 millions
possible scenarios in which at least one component diverge from the specification.

\section{Discussions}\label{sec:discussion}
\paragraph{Abstraction Level}
The level of abstraction is always a concern when it comes to the definition of
a system model.  In this paper, we justified our assumptions on the atomicity
of ports, channels, and functional blocks but nothing prevents a modeler to
consider other atomic element of the system specification and design.  Those
additional elements may detail the system and may change the syntactic sugar we
defined; however, it won't change the underlying (semantic) reasoning unless
another calculus is used (instead of RCC).  Furthermore, other Infosec
properties may be considered, such as authenticity or non-repudiation. We,
however, leave them as future development.

\paragraph{$\abftheory$ Falsification} The theory we developed can be falsified
by providing an experiment that shows a weaknesses not predicted by the
$\abftheory$-theory, or by showing that the weaknesses predicted for a certain
experiment are impossible to realize.  This theory, however, needs to be
detailed to be applicable to a full engineering process (we only cover the
first steps).  We propose the following complete process as a draft for
potential next steps in this direction.
\begin{enumerate}
\item System Specification
\item Architecture Design
\item Security Risk Assessment and Security Requirement Extraction
\item Design Refinement, where the semantics of the functional behavior and
	communication protocols are defined 
\item Security Requirement Extension, based on the design refinement
\item Verification of Security Requirements (data flow, behaviors, \& protocol
	logic)
\item Software Prototype Implementation/Synthesis
\item Automated Security Testing (abstract test-suite generation \&
	concretization)
\item Hardware Prototype Implementation
\item Automated Security Testing
\item Penetration Testing
  \end{enumerate}

\paragraph{Relation with DY}
The DY attacker model has been extensively
used for the identification of logical weaknesses in security protocols.
Our theory should predict all possible weaknesses which, in turn,
should enable all the attacks identified by the DY model. 
A proof that the two theories/models correctly predicts
(or lacks in predicting) weaknesses and related protocol flaws would
connect risk assessment and protocol verification, the first
two steps of the secure engineering of systems.

\section{Conclusion and Future Works}
We proposed a foundational theory on security, arguing that
security-related issues are not related to the maliciousness of an agent but to
the vagueness of the security controls on the engineering processes.
The current state-of-the-art processes allows, given a specification,
an engineer to design the system in such a way that security issues arise due to the lack
of proper security risk assessment processes.
Those design, again, lack of security verification processes based on a solid 
security foundational theory and then permit the generation of insecure implementation.
The security verification and test-case generation will be the focus of our next steps.

We conclude that the problem of security is a problem related to the 
many possible design and, in turn, implementation given a specification.
The problem is correlated to the epistemological search for truth,
where the challenge is to relate Information and Belief to Human Knowledge.
Scientifically, generalizing the Human to an Agent, the problem is
to relate Assertions and Behaviors, to Facts. From an engineering standpoint,
by generalizing the concept of Agents as architectural subsystems, the
problem is to link Channels (and Ports) and Functional Architectures to 
Requirements.

\section*{Acknowledgment}
We thank Katia Santac\`a for the help in the development of the initial ideas, and for
the helpful discussions that allowed us to make this step.

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

%\bibliographystyle{IEEEtran}
%\bibliography{../literature}
\printbibliography

\appendix
\begin{sidewaysfigure*}
	\centering
	\includegraphics[width=\textwidth]{secra_classDiagram.png}
	\caption{Security Risk Assessment -- Class Diagram}
	\label{fig:secraclassdiagram}
\end{sidewaysfigure*}

\end{document}