### 2020-06-09 reports/confidential/selectiveinjectionport.tex

\begin{displaymath}
\po{x}{y}=\overlaps{x}{y}\wedge\neg\partof{x}{y}\wedge\neg\partof{y}{x}\\
\end{displaymath}
and
\begin{align*}
\overlaps{x}{y}=\exists z.~ &\partof{z}{x} &\wedge& \partof{z}{y}\\
&\forall z'.~z'\subseteq z\Rightarrow z'\subseteq x && \forall z''.~z''\subseteq z\Rightarrow z'\subseteq y
\end{align*}
and
\begin{equation*}
\begin{split}
\neg\partof{x}{y}=&\neg[\forall\varphi.~(\varphi\subseteq x)\Rightarrow(\varphi\subseteq y)]\\
&\neg[\forall\varphi.~\neg(\varphi\subseteq x)\vee(\varphi\subseteq y)]\\
&\exists\varphi.~(\varphi\subseteq x)\wedge(\varphi\not\subseteq y)
\end{split}
\end{equation*}
Therefore
\begin{displaymath}
\po{\assertionRegion}{\beliefRegion}=(\varphi\subseteq\assertionRegion\wedge\varphi\subseteq\beliefRegion)\wedge(\varphi'\subseteq\assertionRegion\wedge\varphi'\not\subseteq\beliefRegion)\wedge(\varphi''\subseteq\assertionRegion\wedge\varphi''\subseteq\beliefRegion)
\end{displaymath}
Which means that, when a communication over an input-port results in $\po{\assertionRegion}{\beliefRegion}$, the port is:
\begin{itemize}
\item correctly receiving incoming communication as Assertion and sending it to the recipient agent as Beliefs, \emph{and} ($s\rightarrow a$)
\item incorrectly, dropping some input communication (Assertions) ($s\dropport a$), \emph{and}
\item incorrectly, inserting some new Beliefs ($s\insertport a$)
\end{itemize}
Which we abbreviate as a \emph{selective injection} port: $s\selectiveinjectionport a$.